---
title: "Debug PRS Report"
format:
  html:
    toc: true
    embed-resources: true
    self-contained: true
    fig-width: 12
    fig-height: 8
    theme: default
    code-fold: true
    code-tools: false
    fig-dpi: 300
    fig-format: "png"
    fig-retina: 2
execute:
  cache: false
  freeze: auto
params:
  patient_id: "10003"
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
#| label: setup

# Core data manipulation and visualization
library(tidyverse)
library(patchwork)
library(ggpubr)
library(viridis)
library(showtext)
library(knitr)
# if (!requireNamespace("pak", quietly = TRUE)) {
#   install.packages("pak")
# }
# pak:pak("showtext")

# Set up FontAwesome
use_fontawesome <- FALSE
if (file.exists("./fontawesome-webfont.ttf")) {
  font_add('FontAwesome', './fontawesome-webfont.ttf')
  showtext_auto()
  use_fontawesome <- TRUE
  message("FontAwesome font loaded successfully")
} else {
  warning("FontAwesome font not found. Using circle fallback.")
}

# Helper function to create nicer labels
nice_label <- function(x) {
  x %>% 
    str_replace_all("_", " ") %>%
    str_to_title()
}

# Get current date
report_date <- format(Sys.Date(), "%Y-%m-%d")
```

```{r load-data, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
#| label: load-data

# Set paths
score_files <- list.files(".", pattern = "pgs.txt", full.names = FALSE)
popsim_files <- list.files(".", pattern = "popsimilarity.txt", full.names = FALSE)

# Load score data
scores <- read_tsv(file.path(score_files))
popsim <- read_tsv(file.path(popsim_files))

# Create comprehensive data with all necessary variables
scores_popsim <- scores %>%
  left_join(popsim %>% select(IID, MostSimilarPop), by = "IID") %>% 
  mutate(simple_id = str_extract(IID, "[^_]+$"))

# Calculate percentiles for each PRS score separately
scores_popsim <- scores_popsim %>%
  group_by(PGS) %>%
  mutate(Overall_Percentile = round(percent_rank(Z_MostSimilarPop) * 100, 1)) %>%
  group_by(PGS, MostSimilarPop) %>%
  mutate(Population_Percentile = round(percent_rank(Z_MostSimilarPop) * 100, 1)) %>%
  ungroup()

# params$patient_id <- "10003"
patient_id <- "BCIA-01025256"
# Look up full IID from simple ID
full_patient_id <- scores_popsim %>%
  filter(simple_id == patient_id) %>%
  pull(IID) %>%
  first()

if (is.na(full_patient_id)) {
  stop("Patient ID not found")
}

# Get all PRS scores for this patient
patient_pgs_scores <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  pull(PGS) %>%
  unique()

# Get first PRS score for testing
test_pgs_id <- patient_pgs_scores[1]

# Get patient data for testing
patient_data <- scores_popsim %>%
  filter(IID == full_patient_id, PGS == test_pgs_id)

patient_percentile <- patient_data %>% pull(Overall_Percentile)
```

# Debug PRS Report

**Date:** `r report_date`  
**Patient ID:** `r params$patient_id`  
**Test PRS:** `r test_pgs_id`  
**Patient Percentile:** `r patient_percentile`

## Test Bell Curve

### Step 1: Basic Bell Curve Data
```{r bell-data, echo=FALSE}
#| label: bell-data

# Create bell curve data
x <- rnorm(500)
y <- dnorm(x)
df <- data.frame(x=x, y=y)
q <- quantile(df$x)

df_plot <- ggplot(df) + geom_line(aes(x,y)) + 
           scale_x_continuous(breaks= seq(-4,4,0.5)) +
           annotate(geom="text", x=q, y=0, label=names(q)) + 
           theme(text = element_text(size=22))
df_plot
```



### Step 2: Add Risk Zones
```{r bell-zones, echo=FALSE, fig.width=12, fig.height=8}
#| label: bell-zones

# Add risk zones to the working plot
df_plot +
  annotate("rect", xmin = -4, xmax = -1.5, ymin = 0, ymax = 0.4, 
           fill = "lightblue", alpha = 0.3) +
  annotate("rect", xmin = -1.5, xmax = 1.5, ymin = 0, ymax = 0.4, 
           fill = "lightyellow", alpha = 0.3) +
  annotate("rect", xmin = 1.5, xmax = 4, ymin = 0, ymax = 0.4, 
           fill = "lightcoral", alpha = 0.3)
```

### Step 3: Add Patient Marker
```{r bell-patient, echo=FALSE, fig.width=12, fig.height=8}
#| label: bell-patient

# Add patient marker to the working plot
df_plot +
  annotate("rect", xmin = -4, xmax = -1.5, ymin = 0, ymax = 0.4, 
           fill = "lightblue", alpha = 0.3) +
  annotate("rect", xmin = -1.5, xmax = 1.5, ymin = 0, ymax = 0.4, 
           fill = "lightyellow", alpha = 0.3) +
  annotate("rect", xmin = 1.5, xmax = 4, ymin = 0, ymax = 0.4, 
           fill = "lightcoral", alpha = 0.3) +
  geom_vline(xintercept = qnorm(patient_percentile/100), color = "red", linewidth = 2) +
  annotate("text", x = qnorm(patient_percentile/100), y = 0.4, 
           label = paste("Patient:", round(patient_percentile), "th percentile"), 
           color = "red", size = 5, hjust = -0.1)
```

### Step 4: Add Labels and Icons
```{r bell-labels, echo=FALSE, fig.width=16, fig.height=4}
#| label: bell-labels

# Complete bell curve with all elements
df_plot +
  annotate("rect", xmin = -4, xmax = -1.5, ymin = 0, ymax = 0.4, 
           fill = "lightblue", alpha = 0.3) +
  annotate("rect", xmin = -1.5, xmax = 1.5, ymin = 0, ymax = 0.4, 
           fill = "lightyellow", alpha = 0.3) +
  annotate("rect", xmin = 1.5, xmax = 4, ymin = 0, ymax = 0.4, 
           fill = "lightcoral", alpha = 0.3) +
  # Patient line (limited height to match bell curve)
  annotate("segment", x = qnorm(patient_percentile/100), xend = qnorm(patient_percentile/100), 
           y = 0, yend = 0.4, color = "red", linewidth = 2) +
  # Add human icons in each risk zone using FontAwesome - positioned very close to x-axis
  annotate("text", x = -2.5, y = 0.1, label = if(use_fontawesome) "\uf007" else "●", 
           family = if(use_fontawesome) "FontAwesome" else "sans",
           size = 8, color = "blue") +
  annotate("text", x = -0.5, y = 0.1, label = if(use_fontawesome) "\uf007" else "●", 
           family = if(use_fontawesome) "FontAwesome" else "sans",
           size = 8, color = "gray50") +
  annotate("text", x = 0, y = 0.1, label = if(use_fontawesome) "\uf007" else "●", 
           family = if(use_fontawesome) "FontAwesome" else  "sans",
           size = 8, color = "gray50") +
  annotate("text", x = 0.5, y = 0.1, label = if(use_fontawesome) "\uf007" else "●", 
           family = if(use_fontawesome) "FontAwesome" else "sans",
           size = 8, color = "gray50") +
  annotate("text", x = 2.5, y = 0.1, label = if(use_fontawesome) "\uf007" else "●", 
           family = if(use_fontawesome) "FontAwesome" else "sans",
           size = 8, color = "red") +
  # Patient marker icon - positioned at bell curve height
  annotate("text", x = qnorm(patient_percentile/100), y = 0.4, 
           label = if(use_fontawesome) "\uf007" else "●", 
           family = if(use_fontawesome) "FontAwesome" else "sans",
           size = 12, color = "red") +
  # Risk labels - positioned below the icons
  annotate("text", x = -2.5, y = 0.05, label = "Lower Risk", size = 4) +
  annotate("text", x = 0, y = 0.05, label = "Average Risk", size = 4) +
  annotate("text", x = 2.5, y = 0.05, label = "Higher Risk", size = 4) +
  # Text below the red patient marker line
  annotate("text", x = qnorm(patient_percentile/100), y = 0.45, 
           label = paste0("Your polygenic score is in the ", round(patient_percentile), "th percentile"), 
           color = "black", size = 4) +
  # Limit y-axis to include text and remove axis ticks
  scale_y_continuous(limits = c(0, 0.5)) +
  # Publication theme without grids and axis ticks
  theme_pubr(base_size = 16) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5)
  ) +
  labs(
    title = paste("Polygenic Risk Score:", test_pgs_id),
    subtitle = "Normal Distribution with Risk Zones"
  )
```

## Test Horizontal Risk Bar

```{r horizontal-bar-test, echo=FALSE, fig.width=14, fig.height=4}
#| label: horizontal-bar-test

# Create horizontal risk bar visualization
ggplot() +
  # Main horizontal bar with gradient colors
  geom_rect(aes(xmin = 0, xmax = 20, ymin = 0, ymax = 1), 
            fill = "#3498db", alpha = 0.8) +  # Softer blue
  geom_rect(aes(xmin = 20, xmax = 80, ymin = 0, ymax = 1), 
            fill = "#ecf0f1", alpha = 0.8) +  # Light gray
  geom_rect(aes(xmin = 80, xmax = 100, ymin = 0, ymax = 1), 
            fill = "#e74c3c", alpha = 0.8) +  # Softer red
  
  # Human icons on the bar
  # Low risk (1 icon)
  geom_text(aes(x = 10, y = 0.5), label = if(use_fontawesome) "\uf007" else "●", 
            family = if(use_fontawesome) "FontAwesome" else "sans",
            size = 12, color = "#2980b9") +
  
  # Average risk (7 icons)
  geom_text(aes(x = 30, y = 0.5), label = if(use_fontawesome) "\uf007" else "●", 
            family = if(use_fontawesome) "FontAwesome" else "sans",
            size = 12, color = "#7f8c8d") +
  geom_text(aes(x = 40, y = 0.5), label = if(use_fontawesome) "\uf007" else "●", 
            family = if(use_fontawesome) "FontAwesome" else "sans",
            size = 12, color = "#7f8c8d") +
  geom_text(aes(x = 50, y = 0.5), label = if(use_fontawesome) "\uf007" else "●", 
            family = if(use_fontawesome) "FontAwesome" else "sans",
            size = 12, color = "#7f8c8d") +
  geom_text(aes(x = 60, y = 0.5), label = if(use_fontawesome) "\uf007" else "●", 
            family = if(use_fontawesome) "FontAwesome" else "sans",
            size = 12, color = "#7f8c8d") +
  geom_text(aes(x = 70, y = 0.5), label = if(use_fontawesome) "\uf007" else "●", 
            family = if(use_fontawesome) "FontAwesome" else "sans",
            size = 12, color = "#7f8c8d") +
  
  # High risk (1 icon)
  geom_text(aes(x = 90, y = 0.5), label = if(use_fontawesome) "\uf007" else "●", 
            family = if(use_fontawesome) "FontAwesome" else "sans",
            size = 12, color = "#c0392b") +
  
  # Patient marker
  geom_vline(xintercept = patient_percentile, color = "black", linewidth = 2) +
  geom_text(aes(x = patient_percentile, y = 1.5), 
            label = if(use_fontawesome) "\uf007" else "●", 
            family = if(use_fontawesome) "FontAwesome" else "sans",
            size = 16, color = risk_info$color) +
  
  # Risk labels below
  geom_text(aes(x = 10, y = -0.3), label = "Lower risk", 
            size = 10, color = "#2980b9", fontface = "bold") +
  geom_text(aes(x = 50, y = -0.3), label = "Average risk", 
            size = 10, color = "#2c3e50", fontface = "bold") +
  geom_text(aes(x = 90, y = -0.3), label = "Higher risk", 
            size = 10, color = "#c0392b", fontface = "bold") +
  
  # Percentile markers
  geom_text(aes(x = 0, y = -0.6), label = "0", size = 8, fontface = "bold") +
  geom_text(aes(x = 20, y = -0.6), label = "20", size = 8, fontface = "bold") +
  geom_text(aes(x = 50, y = -0.6), label = "50", size = 8, fontface = "bold") +
  geom_text(aes(x = 80, y = -0.6), label = "80", size = 8, fontface = "bold") +
  geom_text(aes(x = 100, y = -0.6), label = "100", size = 8, fontface = "bold") +
  
  # Scales and theme
  scale_x_continuous(limits = c(-5, 105), breaks = c(0, 20, 50, 80, 100)) +
  scale_y_continuous(limits = c(-0.8, 2)) +
  labs(
    title = "Risk Level Distribution",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    plot.title = element_text(size = 24, face = "bold", hjust = 0.5, margin = margin(b = 15)),
    plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
    plot.background = element_rect(fill = "white", color = NA)
  )
```

## Test Grid Risk Visualization

```{r grid-risk-test, echo=FALSE, fig.width=12, fig.height=8}
#| label: grid-risk-test

# Create grid positions (10 rows x 10 columns)
n_rows <- 10
n_cols <- 10
grid_positions <- expand.grid(
  row = 1:n_rows,
  col = 1:n_cols
) %>%
  mutate(
    percentile_range = case_when(
      row == 1 ~ "0-10",
      row == 10 ~ "91-100",
      TRUE ~ paste0(((row - 1) * 10 + 1), "-", (row * 10))
    ),
    row_start = (row - 1) * 10,
    row_end = row * 10,
    is_patient_row = patient_percentile > row_start & patient_percentile <= row_end
  )

# Create the plot
ggplot(grid_positions, aes(x = col, y = row, color = is_patient_row)) +
  annotate(
    "segment",
    x = n_cols + 1.2,
    xend = n_cols + 1.2,
    y = 1,
    yend = n_rows,
    arrow = arrow(length = unit(0.3, "cm"), ends = "last"),
    color = "black"
  ) +
  geom_text(
    label = if(use_fontawesome) "\uf007" else "●",
    family = if(use_fontawesome) "FontAwesome" else "sans",
    size = 16,
    vjust = 0.5,
    hjust = 0.5,
    show.legend = FALSE
  ) +
  scale_color_manual(values = c("FALSE" = "gray80", "TRUE" = "#3366CC")) +
  scale_y_continuous(
    breaks = 1:n_rows,
    labels = c("0-10", "11-20", "21-30", "31-40", "41-50", 
              "51-60", "61-70", "71-80", "81-90", "91-100"),
    expand = expansion(mult = c(0.05, 0.05))
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0.02, 0.15)),
    limits = c(0.5, n_cols + 1.5)
  ) +
  labs(
    title = "Overall Risk Percentile",
    subtitle = paste0(round(patient_percentile, 1), "th percentile"),
    y = "Percentile Range",
    x = NULL
  ) +
  theme_pubr(base_size = 32) +
  theme(
    plot.title = element_text(size = 42, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 36, hjust = 0.5),
    axis.title = element_text(size = 36, face = "bold"),
    axis.text = element_text(size = 30),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
  ) +
  annotate(
    "text",
    x = n_cols + 1.5,
    y = n_rows,
    label = "risque plus élevé",
    hjust = 0,
    vjust = 0.5,
    size = 12
  ) +
  annotate(
    "text",
    x = n_cols + 1.5,
    y = 1,
    label = "risque plus faible",
    hjust = 0,
    vjust = -0.5,
    size = 12
  )
```

## Test Density Plot

```{r density-test, echo=FALSE, fig.width=12, fig.height=8}
#| label: density-test

# Filter data for the test PRS
filtered_data <- scores_popsim %>% filter(PGS == test_pgs_id)

# Get the x-axis limits for arrow placement
max_density <- max(density(filtered_data$Z_MostSimilarPop)$y)

ggplot() +
  geom_density(
    data = filtered_data,
    aes(x = Z_MostSimilarPop),
    fill = "steelblue",
    alpha = 0.5,
    color = "black",
    linewidth = 1.5
  ) +
  geom_vline(
    data = filtered_data %>% filter(IID == full_patient_id),
    aes(xintercept = Z_MostSimilarPop),
    color = "red",
    linetype = "dashed",
    linewidth = 1.5
  ) +
  geom_text(
    data = filtered_data %>% filter(IID == full_patient_id),
    aes(x = Z_MostSimilarPop, y = max_density * 0.9, 
        label = paste("Patient", params$patient_id)),
    color = "red",
    hjust = -0.1,
    vjust = 0,
    size = 12
  ) +
  labs(
    title = "Overall Distribution of PRS Scores",
    x = "Z-Score (Most Similar Population)",
    y = "Density"
  ) +
  theme_pubr(base_size = 28) +
  theme(
    plot.title = element_text(size = 36, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 32, face = "bold"),
    axis.text = element_text(size = 28),
    plot.margin = margin(t = 20, r = 20, b = 20, l = 20)
  )
```

## Test Histogram

```{r histogram-test, echo=FALSE, fig.width=12, fig.height=8}
#| label: histogram-test

# Filter data for the test PRS
filtered_data <- scores_popsim %>% filter(PGS == test_pgs_id)

# Get the x-axis limits for arrow placement
max_density <- max(density(filtered_data$Z_MostSimilarPop)$y)

ggplot(filtered_data, aes(x = Z_MostSimilarPop)) +
  geom_histogram(binwidth = 0.5, fill = "steelblue", alpha = 0.7, color = "black", linewidth = 0.8) +
  # Add vertical line for patient
  geom_vline(
    data = filtered_data %>% filter(IID == full_patient_id),
    aes(xintercept = Z_MostSimilarPop),
    color = "red",
    linetype = "dashed",
    linewidth = 1.5
  ) +
  # Add text annotation for patient
  geom_text(
    data = filtered_data %>% filter(IID == full_patient_id),
    aes(x = Z_MostSimilarPop, y = max_density * 0.8, 
        label = paste("Patient", params$patient_id)),
    color = "red",
    hjust = -0.1,
    vjust = -10,
    size = 12
  ) +
  labs(
    title = "Distribution of PRS Scores",
    subtitle = "Patient highlighted in red",
    x = "Z-Score (Most Similar Population)",
    y = "Frequency"
  ) +
  theme_pubr(base_size = 28) +
  theme(
    plot.title = element_text(size = 36, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 32, hjust = 0.5),
    axis.title = element_text(size = 32, face = "bold"),
    axis.text = element_text(size = 28),
    plot.margin = margin(t = 20, r = 20, b = 20, l = 20)
  )
```

## Test Boxplot

```{r boxplot-test, echo=FALSE, fig.width=12, fig.height=8}
#| label: boxplot-test

# Filter data for the test PRS
filtered_data <- scores_popsim %>% filter(PGS == test_pgs_id)

ggplot(filtered_data, aes(x = PGS, y = Z_MostSimilarPop)) +
  # Add boxplot
  geom_boxplot(alpha = 0.7, width = 0.5, size = 1.5, outlier.size = 3) +
  # Highlight patient point
  geom_point(
    data = filtered_data %>% filter(IID == full_patient_id),
    size = 10,
    shape = 17,
    color = "red"
  ) +
  # Add mean line
  stat_summary(
    fun = mean,
    geom = "crossbar",
    width = 0.5,
    size = 1.2,
    color = "darkblue"
  ) +
  labs(
    title = "Distribution des scores PRS",
    subtitle = paste("Patient", params$patient_id, "en rouge"),
    x = NULL,
    y = "Score PRS (Z_MostSimilarPop)"
  ) +
  theme_pubr(base_size = 28) +
  theme(
    plot.title = element_text(size = 36, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 32, hjust = 0.5),
    axis.title = element_text(size = 32, face = "bold"),
    axis.text = element_text(size = 28)
  )
```

## Patient Summary Table

```{r summary-table, echo=FALSE}
#| label: summary-table

# Display summary for this patient
patient_summary_table <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  select(PGS, Z_MostSimilarPop, Overall_Percentile) %>%
  distinct() %>%
  rename(
    "Score Type" = PGS,
    "Score" = Z_MostSimilarPop,
    "Overall Percentile" = Overall_Percentile
  )

# For HTML output, use kable
knitr::kable(patient_summary_table, format = "html", digits = 2, align = "c", 
             caption = paste("Résumé de tous les scores PRS pour le patient", params$patient_id))
```

## Explanation Text

**Explication du score polygénique :** Votre score polygénique se situe au **`r patient_percentile`e percentile**. Cela ne signifie pas que vous avez une probabilité de `r patient_percentile`% de développer cette condition. Cela signifie plutôt que sur 100 personnes, votre score polygénique est plus élevé que celui de `r patient_percentile` personnes et identique ou inférieur à celui de `r 100 - patient_percentile` personnes. 