---
title: "PRS Report"
format:
  html:
    toc: true  # Add table of contents for multiple scores
    embed-resources: true  # Make self-contained HTML with all resources embedded
    self-contained: true  # Ensure all resources are embedded
    fig-width: 12  # Increased from 8
    fig-height: 8  # Increased from 6
    theme: default
    code-fold: true
    code-tools: false
    fig-dpi: 300
    fig-format: "png"  # Added to ensure consistent rendering
    fig-retina: 2  # Added for better resolution

execute:
  cache: false
  freeze: auto
params:
  patient_id: "default"
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
#| label: setup

# Core data manipulation and visualization
library(tidyverse)  # Includes dplyr, ggplot2, etc.
library(patchwork)  # For combining plots
library(ggpubr)     # For publication-ready plots
library(viridis)    # For color scales

# Text and formatting
library(showtext)   # For custom fonts
library(knitr)      # For table formatting

# Set up FontAwesome
# Check if fonts directory exists directly or in assets/fonts
if (dir.exists("fonts")) {
  font_add('FontAwesome', 'fonts/fontawesome-webfont.ttf')
} else if (dir.exists("assets/fonts")) {
  font_add('FontAwesome', 'assets/fonts/fontawesome-webfont.ttf')
} else {
  warning("Font directory not found. FontAwesome not available.")
}
showtext_auto()

# Helper function to create nicer labels
nice_label <- function(x) {
  x %>% 
    str_replace_all("_", " ") %>%
    str_to_title()
}

# Create a publication-ready theme with scaled elements
publication_theme <- function(base_size = 24) {
  theme_pubr(base_size = base_size) +
    theme(
      plot.title = element_text(size = rel(1.5), face = "bold"),
      plot.subtitle = element_text(size = rel(1.2)),
      axis.title = element_text(size = rel(1.1), face = "bold"),
      axis.text = element_text(size = rel(1)),
      legend.title = element_text(size = rel(1.1)),
      legend.text = element_text(size = rel(1)),
      legend.key.size = unit(1.5, "line"),
      plot.margin = margin(t = 15, r = 15, b = 15, l = 15)
    )
}

# Get current date in a nice format
report_date <- format(Sys.Date(), "%Y-%m-%d")


```

```{r load-data, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
#| label: load-data

# test params$patient_id
# assing 10003_10003 as test
# params = list(patient_id = "10003")

# Set paths
score_files <- list.files(".", pattern = "pgs.txt", full.names = FALSE)
popsim_files <- list.files(".", pattern = "popsimilarity.txt", full.names = FALSE)

# Load score data
scores <- read_tsv(file.path(score_files))

# Load popsim data
popsim <- read_tsv(file.path(popsim_files))

# Get basic stats
n_scores <- length(unique(scores$PGS))
n_samples <- length(unique(scores$IID))

# Create comprehensive run_patient_data with all necessary variables
scores_popsim <- scores %>%
  left_join(popsim %>% select(IID, MostSimilarPop), by = "IID") %>% 
  # Create a simple ID column using the last part after underscore
  mutate(simple_id = str_extract(IID, "[^_]+$"))

# Calculate percentiles for each PGS score separately
scores_popsim <- scores_popsim %>%
  group_by(PGS) %>%
  mutate(Overall_Percentile = round(percent_rank(Z_MostSimilarPop) * 100, 1)) %>%
  group_by(PGS, MostSimilarPop) %>%
  mutate(Population_Percentile = round(percent_rank(Z_MostSimilarPop) * 100, 1)) %>%
  ungroup()

run_patient_data <- scores_popsim %>%
  filter(sampleset != "reference") %>%
  head(64) # FIXME: should not matter, no more than 64 patients in the reference set should be present

# Get reference data
reference_data <- scores_popsim %>% filter(sampleset == "reference")

# Look up full IID from simple ID
full_patient_id <- scores_popsim %>%
  filter(simple_id == params$patient_id) %>%
  pull(IID) %>%
  first()

if (is.na(full_patient_id)) {
  stop("Patient ID not found")
}

# Get all PGS scores for this patient
patient_pgs_scores <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  pull(PGS) %>%
  unique()

# Save summary data to CSV
patient_summary_all <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  select(IID, PGS, Z_MostSimilarPop, Overall_Percentile, Population_Percentile)

write_csv(patient_summary_all, "patient_summaries.csv")

# Get list of PGS scores for this patient
patient_pgs_list <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  pull(PGS) %>%
  unique()

# Create bulleted list of PGS scores
pgs_bullet_list <- paste0("- **", patient_pgs_list, "**", collapse = "\n")
```


```{r functions, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
#| label: functions

# Add a direct scaling function for plots
scale_plot <- function(p) {
  p + theme(
    text = element_text(size = rel(3)),  # Scale all text by 3x
    plot.title = element_text(size = rel(3.5)),
    plot.subtitle = element_text(size = rel(3)),
    axis.title = element_text(size = rel(3)),
    axis.text = element_text(size = rel(2.5)),
    legend.text = element_text(size = rel(2.5)),
    legend.title = element_text(size = rel(3))
  )
}

# Function to create patient risk visualization with human icons - SIMPLIFIED to only show overall percentile
create_patient_risk_viz <- function(patient_id, pgs_id = NULL) {
  # Get patient data, filtered by PGS if provided
  filtered_data <- if (!is.null(pgs_id)) {
    scores_popsim %>% filter(PGS == pgs_id)
  } else {
    scores_popsim
  }
  
  if(nrow(filtered_data %>% filter(IID == full_patient_id)) == 0) {
    return(ggplot() + annotate("text", x = 0, y = 0, label = "Patient not found") + 
           theme_void())
  }

  # Get overall percentile only
  patient_percentile <- filtered_data %>%
    filter(IID == full_patient_id) %>%
    pull(Overall_Percentile)
  
  # Create grid positions (10 rows x 10 columns)
  n_rows <- 10
  n_cols <- 10
  grid_positions <- expand.grid(
    row = 1:n_rows,
    col = 1:n_cols
  ) %>%
    mutate(
      percentile_range = case_when(
        row == 1 ~ "0-10",
        row == 10 ~ "91-100",
        TRUE ~ paste0(((row - 1) * 10 + 1), "-", (row * 10))
      ),
      row_start = (row - 1) * 10,
      row_end = row * 10,
      is_patient_row = patient_percentile > row_start & patient_percentile <= row_end
    )
  
  # Create the plot with publication-ready theme
  ggplot(grid_positions, aes(x = col, y = row, color = is_patient_row)) +
    annotate(
      "segment",
      x = n_cols + 1.2,
      xend = n_cols + 1.2,
      y = 1,
      yend = n_rows,
      arrow = arrow(length = unit(0.3, "cm"), ends = "last"),
      color = "black"
    ) +
    geom_text(
      label = "\uf007",
      family = "FontAwesome",
      size = 16,
      vjust = 0.5,
      hjust = 0.5,
      show.legend = FALSE
    ) +
    scale_color_manual(values = c("FALSE" = "gray80", "TRUE" = "#3366CC")) +
    scale_y_continuous(
      breaks = 1:n_rows,
      labels = c("0-10", "11-20", "21-30", "31-40", "41-50", 
                "51-60", "61-70", "71-80", "81-90", "91-100"),
      expand = expansion(mult = c(0.05, 0.05))
    ) +
    scale_x_continuous(
      expand = expansion(mult = c(0.02, 0.15)),
      limits = c(0.5, n_cols + 1.5)
    ) +
    labs(
      title = "Overall Risk Percentile",
      subtitle = paste0(round(patient_percentile, 1), "th percentile"),
      y = "Percentile Range",
      x = NULL
    ) +
    theme_pubr(base_size = 32) +  # Using theme_pubr directly with large base size
    theme(
      plot.title = element_text(size = 42, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 36, hjust = 0.5),
      axis.title = element_text(size = 36, face = "bold"),
      axis.text = element_text(size = 30),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank()
    ) +
    annotate(
      "text",
      x = n_cols + 1.5,
      y = n_rows,
      label = "higher risk",
      hjust = 0,
      vjust = 0.5,
      size = 12
    ) +
    annotate(
      "text",
      x = n_cols + 1.5,
      y = 1,
      label = "lower risk",
      hjust = 0,
      vjust = -0.5,
      size = 12
    ) 
}

# Function to create density plot for a patient - SIMPLIFIED to only show overall density
create_density_plot <- function(patient_id, pgs_id = NULL) {
  # Filter data by PGS if provided
  filtered_data <- if (!is.null(pgs_id)) {
    scores_popsim %>% filter(PGS == pgs_id)
  } else {
    scores_popsim
  }
  
  # Get the x-axis limits for arrow placement
  max_density <- max(density(filtered_data$Z_MostSimilarPop)$y)
  
  ggplot() +
    geom_density(
      data = filtered_data,
      aes(x = Z_MostSimilarPop),
      fill = "steelblue",
      alpha = 0.5,
      color = "black",
      linewidth = 1.5
    ) +
    geom_vline(
      data = filtered_data %>% filter(IID == full_patient_id),
      aes(xintercept = Z_MostSimilarPop),
      color = "red",
      linetype = "dashed",
      linewidth = 1.5
    ) +
    geom_text(
      data = filtered_data %>% filter(IID == full_patient_id),
      aes(x = Z_MostSimilarPop, y = max_density * 0.9, 
          label = paste("Patient", params$patient_id)),
      color = "red",
      hjust = -0.1,
      vjust = 0,
      size = 12
    ) +
    labs(
      title = "Overall Distribution of PGS Scores",
      x = "Z-Score (Most Similar Population)",
      y = "Density"
    ) +
    theme_pubr(base_size = 28) +  # Consistent use of theme_pubr
    theme(
      plot.title = element_text(size = 36, face = "bold", hjust = 0.5),
      axis.title = element_text(size = 32, face = "bold"),
      axis.text = element_text(size = 28),
      plot.margin = margin(t = 20, r = 20, b = 20, l = 20)
    )
}

# function to create a histogram of the PGS score
create_histogram_density_plot <- function(patient_id, pgs_id = NULL) {
  # Filter data by PGS if provided
  filtered_data <- if (!is.null(pgs_id)) {
    scores_popsim %>% filter(PGS == pgs_id)
  } else {
    scores_popsim
  }
  
  # Get the x-axis limits for arrow placement
  max_density <- max(density(filtered_data$Z_MostSimilarPop)$y)
  
  ggplot(filtered_data, aes(x = Z_MostSimilarPop)) +
    geom_histogram(binwidth = 0.5, fill = "steelblue", alpha = 0.7, color = "black", linewidth = 0.8) +
    # Add vertical line for patient
    geom_vline(
      data = filtered_data %>% filter(IID == full_patient_id),
      aes(xintercept = Z_MostSimilarPop),
      color = "red",
      linetype = "dashed",
      linewidth = 1.5
    ) +
    # Add text annotation for patient
    geom_text(
      data = filtered_data %>% filter(IID == full_patient_id),
      aes(x = Z_MostSimilarPop, y = max_density * 0.8, 
          label = paste("Patient", patient_id)),
      color = "red",
      hjust = -0.1,
      vjust = -10,
      size = 12
    ) +
    labs(
      title = "Distribution of PGS Scores",
      subtitle = "Patient highlighted in red",
      x = "Z-Score (Most Similar Population)",
      y = "Frequency"
    ) +
    theme_pubr(base_size = 28) +  # Using theme_pubr consistently
    theme(
      plot.title = element_text(size = 36, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 32, hjust = 0.5),
      axis.title = element_text(size = 32, face = "bold"),
      axis.text = element_text(size = 28),
      plot.margin = margin(t = 20, r = 20, b = 20, l = 20)
    )
}

# boxplot of the PGS score with patient placement
create_boxplot <- function(patient_id, pgs_id = NULL) {
  # Filter data by PGS if provided
  filtered_data <- if (!is.null(pgs_id)) {
    scores_popsim %>% filter(PGS == pgs_id)
  } else {
    scores_popsim
  }
  
  ggplot(filtered_data, aes(x = PGS, y = Z_MostSimilarPop)) +
    # Add boxplot
    geom_boxplot(alpha = 0.7, width = 0.5, size = 1.5, outlier.size = 3) +
    # Highlight patient point
    geom_point(
      data = filtered_data %>% filter(IID == full_patient_id),
      size = 10,
      shape = 17,
      color = "red"
    ) +
    # Add mean line
    stat_summary(
      fun = mean,
      geom = "crossbar",
      width = 0.5,
      size = 1.2,
      color = "darkblue"
    ) +
    labs(
      title = "PGS Score Distribution",
      subtitle = paste("Patient", patient_id, "highlighted in red"),
      x = NULL,
      y = "PGS Score (Z_MostSimilarPop)"
    ) +
    theme_pubr(base_size = 28) +  # Using theme_pubr consistently
    theme(
      plot.title = element_text(size = 36, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 32, hjust = 0.5),
      axis.title = element_text(size = 32, face = "bold"),
      axis.text = element_text(size = 28)
    )
}
```



# Overall Summary
**Date d'émission :** `r report_date`

::: {.callout-note}
## Ce rapport contient l'analyse des scores de risque polygénique (PRS) pour le patient **`r params$patient_id`**.   

Les poids dérivés des scores polygéniques suivants ont été utilisés dans ce rapport :

`r pgs_bullet_list`

Ces scores peuvent être consultés dans l'entrée correspondante du [Polygenic Score (PGS) Catalog](https://www.pgscatalog.org/): **https://www.pgscatalog.org/**
:::


::: {.callout-tip}
## Interprétation des scores PGS

Pour chaque score polygénique évalué, le profil génétique du patient est analysé en comparaison avec un groupe de référence génétiquement diversifié. Le percentile de risque final est déterminé en alignant le score individuel avec la population de référence présentant l'ascendance génétique la plus proche.
Il est donc essentiel de se référer aux spécifications détaillées disponibles dans le catalogue PGS pour chaque score. La pertinence des estimations de risque pour d'autres populations est directement liée à leur ressemblance avec les cohortes initiales ayant servi à l'élaboration du modèle.
:::


::: {.callout-warning}
## À des fins de recherche uniquement

Une validation par un laboratoire spécialisé en génétique et l'approbation du médecin traitant sont nécessaires avant toute décision médicale ou autre basée sur ces résultats. Ces analyses constituent des données de recherche préliminaires.
:::

::: {.callout-important}
## Contraintes des PRS

Les principales restrictions comprennent une héritabilité incomplète et une connaissance limitée des facteurs de risque non génétiques. Les prédictions comportent une marge d'incertitude naturelle, et le niveau de risque peut varier selon divers éléments comme l'âge ou les traitements. Les scores de risque polygénique servent d'indicateurs prédictifs et ne constituent pas un outil diagnostique.
:::


```{r}
#| label: overall-summary
#| echo: false
#| warning: false
#| message: false

# Display summary of all PGS scores for this patient
patient_summary_table <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  select(PGS, Z_MostSimilarPop, Overall_Percentile) %>%
  distinct() %>%  # remove any duplicate rows
  rename(
    "Score Type" = PGS,
    "Score" = Z_MostSimilarPop,
    "Overall Percentile" = Overall_Percentile
  )

# For HTML output, use kable
knitr::kable(patient_summary_table, format = "html", digits = 2, align = "c", 
             caption = paste("Summary of all PGS scores for Patient", params$patient_id))
```

```{r generate-sections, results='asis', echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8, out.width="100%"}
#| label: generate-sections
#| fig-width: 12
#| fig-height: 8
#| out-width: "100%"
#| fig-dpi: 300

# Loop through each PGS score and create a section for it
for (pgs_id in patient_pgs_scores) {
  # Create section header
  cat("\n\n## Analysis for PGS Score:", pgs_id, "\n\n")
 
  # Risk visualization section
  cat("\n\n### Risk Visualization for", pgs_id, "\n\n")
  
  # Create risk visualization plot only showing overall risk
  risk_plot <- create_patient_risk_viz(params$patient_id, pgs_id)
  
  # Print the risk plot
  print(risk_plot)
  
  # Score distribution section
  cat("\n\n### Score Distribution for", pgs_id, "\n\n")
  
  # Create and print histogram 
  cat("\n\n#### Overall score histogram\n\n")
  hist_plot <- create_histogram_density_plot(params$patient_id, pgs_id)
  print(hist_plot)
  
  # Add risk direction text with arrow 
  cat('<div style="text-align: center; font-size: 24px; margin-top: -20px; margin-bottom: 20px;"><strong>Lower Risk ←———→ Higher Risk</strong></div>\n\n')
  
  # Create and print density plot
  cat("\n\n#### Overall score distribution\n\n")
  density_plot <- create_density_plot(params$patient_id, pgs_id)
  print(density_plot)
  
  # Add risk direction text with arrow b
  cat('<div style="text-align: center; font-size: 24px; margin-top: -20px; margin-bottom: 20px;"><strong>Lower Risk ←———→ Higher Risk</strong></div>\n\n')
}
``` 

