---
title: "PRS Rapport"
format:
  html:
    toc: true  # Add table of contents for multiple scores
    embed-resources: true  # Make self-contained HTML with all resources embedded
    self-contained: true  # Ensure all resources are embedded
    fig-width: 12  # Increased from 8
    fig-height: 8  # Increased from 6
    theme: default
    code-fold: true
    code-tools: false
    fig-dpi: 300
    fig-format: "png"  # Added to ensure consistent rendering
    fig-retina: 2  # Added for better resolution

execute:
  cache: false
  freeze: auto
params:
  patient_id: "default"
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}

# Core data manipulation and visualization
library(tidyverse)  # Includes dplyr, ggplot2, etc.
library(patchwork)  # For combining plots
library(ggpubr)     # For publication-ready plots
library(viridis)    # For color scales

# Text and formatting
library(showtext)   # For custom fonts
library(knitr)      # For table formatting

# Set up FontAwesome
# Check if fonts exists in the same directory as the template
use_fontawesome <- FALSE
if (file.exists("./fontawesome-webfont.ttf")) {
  font_add('FontAwesome', './fontawesome-webfont.ttf')
  showtext_auto()
  use_fontawesome <- TRUE
  message("FontAwesome font loaded successfully")
} else {
  warning("FontAwesome font not found. Using circle fallback.")
  showtext_auto()
}

# Helper function to create nicer labels
nice_label <- function(x) {
  x %>% 
    str_replace_all("_", " ") %>%
    str_to_title()
}

# Create a publication-ready theme with scaled elements
publication_theme <- function(base_size = 24) {
  theme_pubr(base_size = base_size) +
    theme(
      plot.title = element_text(size = rel(1.5), face = "bold"),
      plot.subtitle = element_text(size = rel(1.2)),
      axis.title = element_text(size = rel(1.1), face = "bold"),
      axis.text = element_text(size = rel(1)),
      legend.title = element_text(size = rel(1.1)),
      legend.text = element_text(size = rel(1)),
      legend.key.size = unit(1.5, "line"),
      plot.margin = margin(t = 15, r = 15, b = 15, l = 15)
    )
}

# Get current date in a nice format
report_date <- format(Sys.Date(), "%Y-%m-%d")


```

```{r load-data, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}

# test params$patient_id
# assing 10003_10003 as test
# params = list(patient_id = "10003")

# Set paths
score_files <- list.files(".", pattern = "pgs.txt", full.names = FALSE)
popsim_files <- list.files(".", pattern = "popsimilarity.txt", full.names = FALSE)

# Load score data
scores <- read_tsv(file.path(score_files))

# Load popsim data
popsim <- read_tsv(file.path(popsim_files))

# Get basic stats
n_scores <- length(unique(scores$PGS))
n_samples <- length(unique(scores$IID))

# Create comprehensive run_patient_data with all necessary variables
scores_popsim <- scores %>%
  left_join(popsim %>% select(IID, MostSimilarPop), by = "IID") %>% 
  # Create a simple ID column using the last part after underscore
  mutate(simple_id = str_extract(IID, "[^_]+$"))

# Calculate percentiles for each PRS score separately
scores_popsim <- scores_popsim %>%
  group_by(PGS) %>%
  mutate(Overall_Percentile = round(percent_rank(Z_MostSimilarPop) * 100, 1)) %>%
  group_by(PGS, MostSimilarPop) %>%
  mutate(Population_Percentile = round(percent_rank(Z_MostSimilarPop) * 100, 1)) %>%
  ungroup()

run_patient_data <- scores_popsim %>%
  filter(sampleset != "reference") %>%
  head(64) # FIXME: should not matter, no more than 64 patients in the reference set should be present

# Get reference data
reference_data <- scores_popsim %>% filter(sampleset == "reference")

# Look up full IID from simple ID
full_patient_id <- scores_popsim %>%
  filter(simple_id == params$patient_id) %>%
  pull(IID) %>%
  first()

if (is.na(full_patient_id)) {
  stop("Patient ID not found")
}

# Get all PRS scores for this patient
patient_pgs_scores <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  pull(PGS) %>%
  unique()

# Save summary data to CSV
patient_summary_all <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  select(IID, PGS, Z_MostSimilarPop, Overall_Percentile, Population_Percentile)

write_csv(patient_summary_all, "patient_summaries.csv")

# Get list of PRS scores for this patient
patient_pgs_list <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  pull(PGS) %>%
  unique()

# Create bulleted list of PRS scores
pgs_bullet_list <- paste0("- **", patient_pgs_list, "**", collapse = "\n")
```


```{r functions, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}

# Add a direct scaling function for plots
scale_plot <- function(p) {
  p + theme(
    text = element_text(size = rel(3)),  # Scale all text by 3x
    plot.title = element_text(size = rel(3.5)),
    plot.subtitle = element_text(size = rel(3)),
    axis.title = element_text(size = rel(3)),
    axis.text = element_text(size = rel(2.5)),
    legend.text = element_text(size = rel(2.5)),
    legend.title = element_text(size = rel(3))
  )
}

create_bell_curve_risk_viz <- function(patient_id, pgs_id = NULL) {
  # Restore original visual but ensure we always build on a ggplot object
  filtered_data <- if (!is.null(pgs_id)) {
    scores_popsim %>% dplyr::filter(PGS == pgs_id)
  } else {
    scores_popsim
  }

  if (nrow(filtered_data %>% dplyr::filter(IID == full_patient_id)) == 0) {
    return(ggplot() + annotate("text", x = 0, y = 0, label = "Patient not found") + theme_void())
  }

  patient_percentile <- filtered_data %>%
    dplyr::filter(IID == full_patient_id) %>%
    dplyr::pull(Overall_Percentile)

  # Build bell curve on a fixed grid for stable gradient
  x_vals <- seq(-4, 4, length.out = 2001)
  y_vals <- dnorm(x_vals)
  df_bell <- tibble(x = x_vals, y = y_vals)
  max_y <- max(df_bell$y)
  patient_x <- qnorm(patient_percentile/100)
  # pastel line color that contrasts with all three panels
  patient_line_color <- "#8B5CF6"  # pastel violet
  
  # Percentile thresholds (20th and 80th)
  t_low <- qnorm(0.20)
  t_high <- qnorm(0.80)
  
  p <- ggplot(df_bell, aes(x = x, y = y)) +
    # Create smooth gradient using many small segments
    geom_segment(aes(x = x, xend = x, y = 0, yend = y, color = x), 
                 size = 2.5, alpha = 0.9) +
    scale_color_gradient2(
      low = "#87CEEB",      # sky blue - more pronounced
      mid = "#F0E68C",      # khaki - more pronounced
      high = "#FFB6C1",     # light pink - more pronounced
      midpoint = 0,         # center the gradient at x=0
      guide = "none"
    ) +
    # Bell curve outline
    geom_line(color = "gray70", linewidth = 3, alpha = 0.9) +
    # Visual thresholds (commented out - using gradient colors instead)
    # geom_vline(xintercept = c(t_low, t_high), linetype = "dashed",
    #            linewidth = 1.2, color = c("#5DADE2", "#E57373"), alpha = 0.6) +
    
    # Patient line (limited height to match bell curve)
    annotate("segment", x = patient_x, xend = patient_x, 
             y = 0, yend = max_y, color = patient_line_color, linewidth = 2) +
    
    # Add human icons in each risk zone using FontAwesome - positioned very close to x-axis
    annotate("text", x = -2.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 22, color = "#87CEEB") +  # sky blue - same as gradient
    annotate("text", x = -0.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 18, color = "gray50") +
    annotate("text", x = 0, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 18, color = "gray50") +
    annotate("text", x = 0.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 18, color = "gray50") +
    annotate("text", x = 2.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 22, color = "#FFB6C1") +  # light pink - same as gradient
    
    # Patient marker icon - positioned at bell curve height
    annotate("text", x = patient_x, y = max_y, 
             label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 22, color = patient_line_color) +
    # show patient id next to the marker line
    annotate("text", x = patient_x + 0.05, y = max_y*0.92, label = patient_id,
             color = "black", size = 9, hjust = 0, vjust = 0.5) +
    
    # Risk labels - positioned below the icons
    annotate("text", x = -2.5, y = max_y*0.12, label = "Lower Risk", size = 9) +
    annotate("text", x = 0, y = max_y*0.12, label = "Average Risk", size = 9) +
    annotate("text", x = 2.5, y = max_y*0.12, label = "Higher Risk", size = 9) +
    
    # Text below the red patient marker line
    annotate("text", x = qnorm(patient_percentile/100), y = 0.45, 
             label = paste0("Your polygenic score is in the ", round(patient_percentile), "th percentile"), 
             color = "black", size = 4) +
    
    # Limit y-axis to include text and remove axis ticks
    scale_y_continuous(limits = c(0, max_y), expand = expansion(mult = c(0, 0))) +
    
    # Publication theme without grids and axis ticks
    theme_pubr(base_size = 36) +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.title = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      plot.title = element_text(size = 44, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 32, hjust = 0.5)
    ) +
    labs(
      title = paste("Polygenic Risk Score:", if(!is.null(pgs_id)) pgs_id else "Overall"),
      subtitle = "Normal Distribution with Risk Zones"
    ) +
    # Labels for risk sections (single line to avoid odd breaks)
    annotate("text", x = -2.75, y = max_y*0.55, 
             label = "Fewer individuals have decreased risk", 
             size = 10, color = "#4575b4") +
    annotate("text", x = 0, y = max_y*0.55, 
             label = "Most individuals have average risk", 
             size = 10, color = "gray30") +
    annotate("text", x = 2.75, y = max_y*0.55, 
             label = "Fewer individuals have increased risk", 
             size = 10, color = "#d73027") +
    # Quantile-based x-axis; keep y scale from earlier (0..0.5)
    scale_x_continuous(
      limits = c(-4, 4),
      breaks = qnorm(c(0, 0.05, 0.20, 0.50, 0.80, 0.95, 0.99)),
      labels = c("0", "5th", "20th", "50th", "80th", "95th", "99th")
    )

  return(p)
}

# Function to create patient risk visualization with human icons - only show overall percentile
create_patient_risk_viz <- function(patient_id, pgs_id = NULL) {
  # Get patient data, filtered by PRS if provided
  filtered_data <- if (!is.null(pgs_id)) {
    scores_popsim %>% filter(PGS == pgs_id)
  } else {
    scores_popsim
  }
  
  if(nrow(filtered_data %>% filter(IID == full_patient_id)) == 0) {
    return(ggplot() + annotate("text", x = 0, y = 0, label = "Patient not found") + 
           theme_void())
  }

  # Get overall percentile only
  patient_percentile <- filtered_data %>%
    filter(IID == full_patient_id) %>%
    pull(Overall_Percentile)
  
  # Create grid positions (10 rows x 10 columns)
  n_rows <- 10
  n_cols <- 10
  grid_positions <- expand.grid(
    row = 1:n_rows,
    col = 1:n_cols
  ) %>%
    mutate(
      percentile_range = case_when(
        row == 1 ~ "0-10",
        row == 10 ~ "91-100",
        TRUE ~ paste0(((row - 1) * 10 + 1), "-", (row * 10))
      ),
      row_start = (row - 1) * 10,
      row_end = row * 10,
      is_patient_row = patient_percentile > row_start & patient_percentile <= row_end
    )
  
  # Create the plot with publication-ready theme
  ggplot(grid_positions, aes(x = col, y = row, color = is_patient_row)) +
    annotate(
      "segment",
      x = n_cols + 1.2,
      xend = n_cols + 1.2,
      y = 1,
      yend = n_rows,
      arrow = arrow(length = unit(0.3, "cm"), ends = "last"),
      color = "black"
    ) +
    geom_text(
      label = if(use_fontawesome) "\uf007" else "●",     # FontAwesome person icon or circle fallback
      family = if(use_fontawesome) "FontAwesome" else "sans",
      size = 16,
      vjust = 0.5,
      hjust = 0.5,
      show.legend = FALSE
    ) +
    scale_color_manual(values = c("FALSE" = "gray80", "TRUE" = "#3366CC")) +
    scale_y_continuous(
      breaks = 1:n_rows,
      labels = c("0-10", "11-20", "21-30", "31-40", "41-50", 
                "51-60", "61-70", "71-80", "81-90", "91-100"),
      expand = expansion(mult = c(0.05, 0.05))
    ) +
    scale_x_continuous(
      expand = expansion(mult = c(0.02, 0.15)),
      limits = c(0.5, n_cols + 1.5)
    ) +
    labs(
      title = "Overall Risk Percentile",
      subtitle = paste0(round(patient_percentile, 1), "th percentile"),
      y = "Percentile Range",
      x = NULL
    ) +
    theme_pubr(base_size = 32) +  # Using theme_pubr directly with large base size
    theme(
      plot.title = element_text(size = 42, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 36, hjust = 0.5),
      axis.title = element_text(size = 36, face = "bold"),
      axis.text = element_text(size = 30),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank()
    ) +
    annotate(
      "text",
      x = n_cols + 1.5,
      y = n_rows,
      label = "risque plus élevé",
      hjust = 0,
      vjust = 0.5,
      size = 12
    ) +
    annotate(
      "text",
      x = n_cols + 1.5,
      y = 1,
      label = "risque plus faible",
      hjust = 0,
      vjust = -0.5,
      size = 12
    ) 
}

# Function to create density plot for a patient - SIMPLIFIED to only show overall density
create_density_plot <- function(patient_id, pgs_id = NULL) {
  # Filter data by PRS if provided
  filtered_data <- if (!is.null(pgs_id)) {
    scores_popsim %>% filter(PGS == pgs_id)
  } else {
    scores_popsim
  }
  
  # Get the x-axis limits for arrow placement
  max_density <- max(density(filtered_data$Z_MostSimilarPop)$y)
  
  ggplot() +
    geom_density(
      data = filtered_data,
      aes(x = Z_MostSimilarPop),
      fill = "steelblue",
      alpha = 0.5,
      color = "black",
      linewidth = 1.5
    ) +
    geom_vline(
      data = filtered_data %>% filter(IID == full_patient_id),
      aes(xintercept = Z_MostSimilarPop),
      color = "red",
      linetype = "dashed",
      linewidth = 1.5
    ) +
    geom_text(
      data = filtered_data %>% filter(IID == full_patient_id),
      aes(x = Z_MostSimilarPop, y = max_density * 0.9, 
          label = paste("Patient", params$patient_id)),
      color = "red",
      hjust = -0.1,
      vjust = 0,
      size = 12
    ) +
    labs(
      title = "Overall Distribution of PRS Scores",
      x = "Z-Score (Most Similar Population)",
      y = "Density"
    ) +
    theme_pubr(base_size = 28) +  # Consistent use of theme_pubr
    theme(
      plot.title = element_text(size = 36, face = "bold", hjust = 0.5),
      axis.title = element_text(size = 32, face = "bold"),
      axis.text = element_text(size = 28),
      plot.margin = margin(t = 20, r = 20, b = 20, l = 20)
    )
}

# function to create a histogram of the PRS score
create_histogram_density_plot <- function(patient_id, pgs_id = NULL) {
  # Filter data by PRS if provided
  filtered_data <- if (!is.null(pgs_id)) {
    scores_popsim %>% filter(PGS == pgs_id)
  } else {
    scores_popsim
  }
  
  # Get the x-axis limits for arrow placement
  max_density <- max(density(filtered_data$Z_MostSimilarPop)$y)
  
  ggplot(filtered_data, aes(x = Z_MostSimilarPop)) +
    geom_histogram(binwidth = 0.5, fill = "steelblue", alpha = 0.7, color = "black", linewidth = 0.8) +
    # Add vertical line for patient
    geom_vline(
      data = filtered_data %>% filter(IID == full_patient_id),
      aes(xintercept = Z_MostSimilarPop),
      color = "red",
      linetype = "dashed",
      linewidth = 1.5
    ) +
    # Add text annotation for patient
    geom_text(
      data = filtered_data %>% filter(IID == full_patient_id),
      aes(x = Z_MostSimilarPop, y = max_density * 0.8, 
          label = paste("Patient", patient_id)),
      color = "red",
      hjust = -0.1,
      vjust = -10,
      size = 12
    ) +
    labs(
      title = "Distribution of PRS Scores",
      subtitle = "Patient highlighted in red",
      x = "Z-Score (Most Similar Population)",
      y = "Frequency"
    ) +
    theme_pubr(base_size = 28) +  # Using theme_pubr consistently
    theme(
      plot.title = element_text(size = 36, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 32, hjust = 0.5),
      axis.title = element_text(size = 32, face = "bold"),
      axis.text = element_text(size = 28),
      plot.margin = margin(t = 20, r = 20, b = 20, l = 20)
    )
}

# boxplot of the PRS score with patient placement
create_boxplot <- function(patient_id, pgs_id = NULL) {
  # Filter data by PRS if provided
  filtered_data <- if (!is.null(pgs_id)) {
    scores_popsim %>% filter(PGS == pgs_id)
  } else {
    scores_popsim
  }
  
  ggplot(filtered_data, aes(x = PGS, y = Z_MostSimilarPop)) +
    # Add boxplot
    geom_boxplot(alpha = 0.7, width = 0.5, size = 1.5, outlier.size = 3) +
    # Highlight patient point
    geom_point(
      data = filtered_data %>% filter(IID == full_patient_id),
      size = 10,
      shape = 17,
      color = "red"
    ) +
    # Add mean line
    stat_summary(
      fun = mean,
      geom = "crossbar",
      width = 0.5,
      size = 1.2,
      color = "darkblue"
    ) +
    labs(
      title = "Distribution des scores PRS",
      subtitle = paste("Patient", patient_id, "en rouge"),
      x = NULL,
      y = "Score PRS (Z_MostSimilarPop)"
    ) +
    theme_pubr(base_size = 28) +  # Using theme_pubr consistently
    theme(
      plot.title = element_text(size = 36, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 32, hjust = 0.5),
      axis.title = element_text(size = 32, face = "bold"),
      axis.text = element_text(size = 28)
    )
}
```



# Résumé général
**Date d'émission :** `r report_date`

::: {.callout-note}
## Ce rapport contient l'analyse des scores de risque polygénique (PRS) pour le patient **`r params$patient_id`**.   

Les poids dérivés des scores polygéniques suivants ont été utilisés dans ce rapport :

`r pgs_bullet_list`

Ces scores peuvent être consultés dans l'entrée correspondante du [Polygenic Score Catalog](https://www.pgscatalog.org/): **https://www.pgscatalog.org/**

Ce score a été ajusté avec le Z-score de la population la plus similaire (Z_MostSimilarPop) en utilisant [pgsc-calc](https://pgsc-calc.readthedocs.io/en/latest/) version 2.0.1 [^1].
:::


::: {.callout-tip}
## Interprétation des scores PRS

Pour chaque score polygénique évalué, le profil génétique du patient est analysé en comparaison avec un groupe de référence génétiquement diversifié. Le percentile de risque final est déterminé en alignant le score individuel avec la population de référence présentant l'ascendance génétique la plus proche.
Il est donc essentiel de se référer aux spécifications détaillées disponibles dans le Polygenic Score Catalog pour chaque score. La pertinence des estimations de risque pour d'autres populations est directement liée à leur ressemblance avec les cohortes initiales ayant servi à l'élaboration du modèle.
:::


::: {.callout-warning}
## À des fins de recherche uniquement

Une validation par un laboratoire spécialisé en génétique et l'approbation du médecin traitant sont nécessaires avant toute décision médicale ou autre basée sur ces résultats. Ces analyses constituent des données de recherche préliminaires.
:::

::: {.callout-important}
## Contraintes des PRS

Les principales restrictions comprennent une héritabilité incomplète et une connaissance limitée des facteurs de risque non génétiques. Les prédictions comportent une marge d'incertitude naturelle, et le niveau de risque peut varier selon divers éléments comme l'âge ou les traitements. Les scores de risque polygénique servent d'indicateurs prédictifs et ne constituent pas un outil diagnostique.
:::

[^1]: Lambert, S.A., Wingfield, B. et al. (2024) Enhancing the Polygenic Score Catalog with tools for score calculation and ancestry normalization. *Nature Genetics*. doi:10.1038/s41588-024-01937-x.

```{r}
#| echo: false
#| warning: false
#| message: false

# Display summary of all PRS scores for this patient
patient_summary_table <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  select(PGS, Z_MostSimilarPop, Overall_Percentile) %>%
  distinct() %>%  # remove any duplicate rows
  rename(
    "Score Type" = PGS,
    "Score" = Z_MostSimilarPop,
    "Overall Percentile" = Overall_Percentile
  )

# For HTML output, use kable
knitr::kable(patient_summary_table, format = "html", digits = 2, align = "c", 
             caption = paste("Résumé de tous les scores PRS pour le patient", params$patient_id))
```

```{r generate-sections, results='asis', echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8, out.width="100%"}
#| fig-width: 12
#| fig-height: 8
#| out-width: "100%"
#| fig-dpi: 300

# Loop through each PRS score and create a section for it
for (pgs_id in patient_pgs_scores) {
  # Create section header
  cat("\n\n## Analyse du score PRS:", pgs_id, "\n\n")
  
  # NEW: Bell curve risk visualization
  cat("\n\n### Évaluation du risque génétique pour", pgs_id, "\n\n")
  
  # Create and print bell curve visualization
  bell_curve_plot <- create_bell_curve_risk_viz(params$patient_id, pgs_id)
  print(bell_curve_plot)
  
  # Add explanation text
  cat("\n\n")
  cat("**Explication du score polygénique :** Votre score polygénique se situe au **", 
      round(scores_popsim %>% filter(IID == full_patient_id, PGS == pgs_id) %>% pull(Overall_Percentile)), 
      "e percentile**. Cela ne signifie pas que vous avez une probabilité de ", 
      round(scores_popsim %>% filter(IID == full_patient_id, PGS == pgs_id) %>% pull(Overall_Percentile)), 
      "% de développer cette condition. Cela signifie plutôt que sur 100 personnes, votre score polygénique est plus élevé que celui de ", 
      round(scores_popsim %>% filter(IID == full_patient_id, PGS == pgs_id) %>% pull(Overall_Percentile)), 
      " personnes et identique ou inférieur à celui de ", 
      100 - round(scores_popsim %>% filter(IID == full_patient_id, PGS == pgs_id) %>% pull(Overall_Percentile)), 
      " personnes.\n\n")
 
  # Risk visualization section
  cat("\n\n### Visualisation du risque pour", pgs_id, "\n\n")
  
  # Create risk visualization plot only showing overall risk
  risk_plot <- create_patient_risk_viz(params$patient_id, pgs_id)
  
  # Print the risk plot
  print(risk_plot)
  
  # Score distribution section
  cat("\n\n### Distribution du score pour", pgs_id, "\n\n")
  
  # Create and print histogram 
  cat("\n\n#### Histogramme du score global\n\n")
  hist_plot <- create_histogram_density_plot(params$patient_id, pgs_id)
  print(hist_plot)
  
  # Add risk direction text with arrow 
  cat('<div style="text-align: center; font-size: 24px; margin-top: -20px; margin-bottom: 20px;"><strong>Risque plus faible ←———→ Risque plus élevé</strong></div>\n\n')
  
  # Create and print density plot
  cat("\n\n#### Distribution du score global\n\n")
  density_plot <- create_density_plot(params$patient_id, pgs_id)
  print(density_plot)
  
  # Add risk direction text with arrow b
  cat('<div style="text-align: center; font-size: 24px; margin-top: -20px; margin-bottom: 20px;"><strong>Risque plus faible ←———→ Risque plus élevé</strong></div>\n\n')
}
``` 

