---
title: "PRS Rapport"
format:
  html:
    toc: true  # Add table of contents for multiple scores
    embed-resources: true  # Make self-contained HTML with all resources embedded
    self-contained: true  # Ensure all resources are embedded
    fig-width: 12  # Increased from 8
    fig-height: 8  # Increased from 6
    theme: default
    code-fold: true
    code-tools: false
    fig-dpi: 300
    fig-format: "png"  # Added to ensure consistent rendering
    fig-retina: 2  # Added for better resolution

execute:
  cache: false
  freeze: auto
params:
  patient_id: "default"
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}

# Core data manipulation and visualization
library(tidyverse)  # Includes dplyr, ggplot2, etc.
library(patchwork)  # For combining plots
library(ggpubr)     # For publication-ready plots
library(viridis)    # For color scales

# Text and formatting
library(showtext)   # For custom fonts
library(knitr)      # For table formatting

# Set up FontAwesome
# Check if fonts exists in the same directory as the template
use_fontawesome <- FALSE
if (file.exists("fontawesome-webfont.ttf")) {
  font_add('FontAwesome', 'fontawesome-webfont.ttf')
  showtext_auto()
  use_fontawesome <- TRUE
  message("FontAwesome font loaded successfully")
} else {
  warning("FontAwesome font not found. Using circle fallback.")
  showtext_auto()
}

# Helper function to create nicer labels
nice_label <- function(x) {
  x %>% 
    str_replace_all("_", " ") %>%
    str_to_title()
}

# Create a publication-ready theme with scaled elements
publication_theme <- function(base_size = 24) {
  theme_pubr(base_size = base_size) +
    theme(
      plot.title = element_text(size = rel(1.5), face = "bold"),
      plot.subtitle = element_text(size = rel(1.2)),
      axis.title = element_text(size = rel(1.1), face = "bold"),
      axis.text = element_text(size = rel(1)),
      legend.title = element_text(size = rel(1.1)),
      legend.text = element_text(size = rel(1)),
      legend.key.size = unit(1.5, "line"),
      plot.margin = margin(t = 15, r = 15, b = 15, l = 15)
    )
}

# Get current date in a nice format
report_date <- format(Sys.Date(), "%Y-%m-%d")


```

```{r load-data, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}

# test params$patient_id
# assing 10003_10003 as test
# params = list(patient_id = "10003")

# Set paths
score_files <- list.files(".", pattern = "pgs.txt", full.names = FALSE)
popsim_files <- list.files(".", pattern = "popsimilarity.txt", full.names = FALSE)
metadata_file <- list.files(".", pattern = "log_scorefiles.json", full.names = FALSE)

# Load score data
scores <- read_tsv(file.path(score_files))

# Load popsim data
popsim <- read_tsv(file.path(popsim_files))

# Load PGS metadata if available
pgs_metadata <- NULL
if (length(metadata_file) > 0) {
  tryCatch({
    pgs_metadata <- jsonlite::fromJSON(metadata_file[1], simplifyVector = FALSE)
    message("Loaded PGS metadata for ", length(pgs_metadata), " scoring files")
  }, error = function(e) {
    message("Could not load PGS metadata: ", e$message)
  })
}

# Get basic stats
n_scores <- length(unique(scores$PGS))
n_samples <- length(unique(scores$IID))

# Create comprehensive run_patient_data with all necessary variables
scores_popsim <- scores %>%
  left_join(popsim %>% select(IID, MostSimilarPop), by = "IID") %>% 
  # Create a simple ID column using the last part after underscore
  mutate(simple_id = str_extract(IID, "[^_]+$"))

# Calculate percentiles for each PRS score separately
scores_popsim <- scores_popsim %>%
  group_by(PGS) %>%
  mutate(Overall_Percentile = round(percent_rank(Z_MostSimilarPop) * 100, 1)) %>%
  group_by(PGS, MostSimilarPop) %>%
  mutate(Population_Percentile = round(percent_rank(Z_MostSimilarPop) * 100, 1)) %>%
  ungroup()

run_patient_data <- scores_popsim %>%
  filter(sampleset != "reference")

# Get reference data
reference_data <- scores_popsim %>% filter(sampleset == "reference")

# Look up full IID from simple ID
full_patient_id <- scores_popsim %>%
  filter(simple_id == params$patient_id) %>%
  pull(IID) %>%
  first()

if (is.na(full_patient_id)) {
  stop("Patient ID not found")
}

# Get all PRS scores for this patient
patient_pgs_scores <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  pull(PGS) %>%
  unique()

# Save summary data to CSV
patient_summary_all <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  select(IID, PGS, Z_MostSimilarPop, Overall_Percentile, Population_Percentile)

write_csv(patient_summary_all, "patient_summaries.csv")

# Get list of PRS scores for this patient
patient_pgs_list <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  pull(PGS) %>%
  unique()

# Function to get PGS metadata
get_pgs_metadata <- function(pgs_id) {
  if (is.null(pgs_metadata)) return(NULL)
  # Normalize to base PGS accession (handles values like PGS005168_hmPOS_GRCh38)
  base_id <- stringr::str_replace(pgs_id, "^(PGS[0-9]+).*", "\\1")
  for (entry in pgs_metadata) {
    if (entry$pgs_id == base_id) {
      return(entry)
    }
  }
  return(NULL)
}

# Create bulleted list of PRS scores with formatted names and metadata
formatted_pgs_list <- c()
for (pgs_id in patient_pgs_list) {
  metadata <- get_pgs_metadata(pgs_id)
  base_id <- str_replace(pgs_id, "^(PGS\\d+).*", "\\1")
  if (!is.null(metadata) && metadata$header$pgs_name != "") {
    formatted_name <- paste0("**", base_id, "** (", metadata$header$pgs_name, ")")
  } else {
    formatted_name <- paste0("**", base_id, "** *[Score personnalisé]*")
  }
  formatted_pgs_list <- c(formatted_pgs_list, formatted_name)
}

pgs_bullet_list <- paste0("- ", formatted_pgs_list, collapse = "\n")
```


```{r functions, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}

# Add a direct scaling function for plots
scale_plot <- function(p) {
  p + theme(
    text = element_text(size = rel(3)),  # Scale all text by 3x
    plot.title = element_text(size = rel(3.5)),
    plot.subtitle = element_text(size = rel(3)),
    axis.title = element_text(size = rel(3)),
    axis.text = element_text(size = rel(2.5)),
    legend.text = element_text(size = rel(2.5)),
    legend.title = element_text(size = rel(3))
  )
}

create_bell_curve_risk_viz <- function(patient_id, pgs_id = NULL) {
  # Restore original visual but ensure we always build on a ggplot object
  filtered_data <- if (!is.null(pgs_id)) {
    scores_popsim %>% dplyr::filter(PGS == pgs_id)
  } else {
    scores_popsim
  }

  if (nrow(filtered_data %>% dplyr::filter(IID == full_patient_id)) == 0) {
    return(ggplot() + annotate("text", x = 0, y = 0, label = "Patient not found") + theme_void())
  }

  patient_percentile <- filtered_data %>%
    dplyr::filter(IID == full_patient_id) %>%
    dplyr::pull(Overall_Percentile)

  # Build bell curve on a fixed grid for stable gradient
  x_vals <- seq(-4, 4, length.out = 2001)
  y_vals <- dnorm(x_vals)
  df_bell <- tibble(x = x_vals, y = y_vals)
  max_y <- max(df_bell$y)
  patient_x <- qnorm(patient_percentile/100)
  # pastel line color that contrasts with all three panels
  patient_line_color <- "#8B5CF6"  # pastel violet
  
  # Percentile thresholds (20th and 80th)
  t_low <- qnorm(0.20)
  t_high <- qnorm(0.80)
  
  p <- ggplot(df_bell, aes(x = x, y = y)) +
    # Create smooth gradient using many small segments
    geom_segment(aes(x = x, xend = x, y = 0, yend = y, color = x), 
                 size = 2.5, alpha = 0.9) +
    scale_color_gradient2(
      low = "#228B22",      # forest green - better low risk color
      mid = "#DDA0DD",      # plum - better intermediate color
      high = "#DC143C",     # crimson - more pronounced high risk
      midpoint = 0,         # center the gradient at x=0
      guide = "none"
    ) +
    # Bell curve outline
    geom_line(color = "gray70", linewidth = 3, alpha = 0.9) +
    # Visual thresholds (commented out - using gradient colors instead)
    # geom_vline(xintercept = c(t_low, t_high), linetype = "dashed",
    #            linewidth = 1.2, color = c("#5DADE2", "#E57373"), alpha = 0.6) +
    
    # Patient line (limited height to match bell curve, no arrow head)
    annotate("segment", x = patient_x, xend = patient_x, 
             y = 0, yend = max_y, color = patient_line_color, linewidth = 3) +
    
    # Add human icons in each risk zone using FontAwesome - positioned very close to x-axis
    annotate("text", x = -2.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 22, color = "#228B22") +  # forest green - same as gradient
    annotate("text", x = -0.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 18, color = "gray50") +
    annotate("text", x = 0, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 18, color = "gray50") +
    annotate("text", x = 0.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 18, color = "gray50") +
    annotate("text", x = 2.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 22, color = "#DC143C") +  # crimson - same as gradient
    
    # Patient marker icon - positioned at bell curve height
    annotate("text", x = patient_x, y = max_y, 
             label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 22, color = patient_line_color) +
    # show "You" next to the marker line
    annotate("text", x = patient_x + 0.05, y = max_y*0.92, label = "You",
             color = "black", size = 12, hjust = 0, vjust = 0.5) +
    
    # Risk labels - positioned below the icons
    annotate("text", x = -2.5, y = max_y*0.12, label = "Lower Risk", size = 9) +
    annotate("text", x = 0, y = max_y*0.12, label = "Average Risk", size = 9) +
    annotate("text", x = 2.5, y = max_y*0.12, label = "Higher Risk", size = 9) +
    
    # Text below the patient marker line - larger text
    annotate("text", x = qnorm(patient_percentile/100), y = 0.45, 
             label = paste0("Your polygenic score is in the ", round(patient_percentile), "th percentile"), 
             color = "black", size = 6) +
    
    # Limit y-axis to include text and remove axis ticks
    scale_y_continuous(limits = c(0, max_y), expand = expansion(mult = c(0, 0))) +
    
    # Publication theme without grids and axis ticks
    theme_pubr(base_size = 36) +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.title = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      plot.title = element_text(size = 44, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 32, hjust = 0.5)
    ) +
    labs(
      title = paste("Polygenic Risk Score:", if(!is.null(pgs_id)) pgs_id else "Overall"),
      subtitle = "Normal Distribution with Risk Zones"
    ) +
    # Labels for risk sections (single line to avoid odd breaks) - larger text
    annotate("text", x = -2.75, y = max_y*0.55, 
             label = "Fewer individuals have decreased risk", 
             size = 12, color = "#228B22") +
    annotate("text", x = 0, y = max_y*0.55, 
             label = "Most individuals have average risk", 
             size = 12, color = "gray30") +
    annotate("text", x = 2.75, y = max_y*0.55, 
             label = "Fewer individuals have increased risk", 
             size = 12, color = "#DC143C") +
    # Quantile-based x-axis; keep y scale from earlier (0..0.5)
    scale_x_continuous(
      limits = c(-4, 4),
      breaks = qnorm(c(0, 0.05, 0.20, 0.50, 0.80, 0.95, 0.99)),
      labels = c("0", "5th", "20th", "50th", "80th", "95th", "99th")
    )

  return(p)
}
```



# Résumé général
**Date d'émission :** `r report_date`

::: {.callout-note}
## Ce rapport contient l'analyse des scores de risque polygénique (PRS) pour le patient **`r params$patient_id`**.   

Les poids dérivés des scores polygéniques suivants ont été utilisés dans ce rapport :

`r pgs_bullet_list`

```{r, echo=FALSE, warning=FALSE}
# Simple outline for custom scores
custom_scores <- patient_pgs_list[!stringr::str_detect(patient_pgs_list, "^PGS[0-9]+")]
if (length(custom_scores) > 0) {
  cat("**Scores personnalisés utilisés:**\n\n")
  for (score in custom_scores) {
    cat("- **", score, "** *[Score personnalisé: Consulter la source]*\n")
  }
  cat("\n")
}
```

```{r, echo=FALSE, warning=FALSE}
# If metadata is available, display a compact table like the Catalog report, but
# restricted to the scores used for this patient
if (!is.null(pgs_metadata)) {
  patient_base_ids <- unique(stringr::str_replace(patient_pgs_list, "^(PGS[0-9]+).*", "\\1"))
  json_list <- pgs_metadata
  # helper extractors
  extract_chr_handle_null <- function(x, field) {
    return(replace(x[[field]], is.null(x[[field]]), ""))
  }
  link_pgscatalog <- function(id, link_type) {
    if (id != "") stringr::str_glue('<a href="https://www.pgscatalog.org/{link_type}/{id}">{id}</a>') else id
  }
  
  add_note <- function(id, note) {
    if (id != "" && note != "") {
      return(stringr::str_glue("{id} <br /> <small>{note}</small>"))
    } else {
      return(id)
    }
  }
  
  link_traits <- function(trait_efo, mapped) {
    if (length(trait_efo) == 0) {
      return("")
    } else {
      return(purrr::map2_chr(trait_efo, mapped, ~ stringr::str_glue('<a href="http://www.ebi.ac.uk/efo/{.x}">{.y}</a>')))
    }
  }
  
  extract_traits <- function(x) {
    trait_efo <- purrr::map(x, ~ extract_chr_handle_null(.x$header, "trait_efo"))
    mapped <- purrr::map(x, ~ extract_chr_handle_null(.x$header, "trait_mapped"))
    trait_display <- purrr::map2(trait_efo, mapped, link_traits)
    mapped_trait_links <- purrr::map_chr(trait_display, ~ paste(.x, collapse = "<br />"))
    reported_traits <- purrr::map(x, ~ extract_chr_handle_null(.x, "trait_reported"))
    purrr::map2_chr(reported_traits, mapped_trait_links, ~ {
      stringr::str_glue("<u>Reported trait:</u> {.x} <br /> <u>Mapped trait(s):</u> {.y}")
    })
  }
  # filter json entries to those used in this patient
  json_list_patient <- purrr::keep(json_list, ~ .x$pgs_id %in% patient_base_ids)
  if (length(json_list_patient) > 0) {
  # Create metadata table for Catalog scores
  catalog_metadata <- tibble::tibble(
    pgs_id = purrr::map_chr(json_list_patient, "pgs_id"),
    pgs_name = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "pgs_name")),
    pgp_id = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "pgp_id")),
    citation = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "citation")),
    trait_display = extract_traits(json_list_patient),
    n_variants = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "variants_number")),
    genome_build = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "genome_build")),
    harmonised_build = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "HmPOS_build"))
  ) %>%
      dplyr::mutate(
        pgs_id = purrr::map2_chr(pgs_id, pgs_name, ~ add_note(link_pgscatalog(.x, "score"), .y)),
        pgp_id = purrr::map2_chr(pgp_id, citation, ~ add_note(link_pgscatalog(.x, "publication"), .y)),
        genome_build = purrr::map2_chr(genome_build, harmonised_build, ~ {
          if (.y != "") {
            stringr::str_glue("<u>Original build:</u> {.x} <br /> <u>Harmonised build:</u> {.y}")
          } else {
            .x
          }
        })
      )
    
    scorefile_metadata <- catalog_metadata
    
    scorefile_metadata <- scorefile_metadata %>%
      dplyr::select(pgs_id, pgp_id, trait_display, n_variants, genome_build) %>%
      dplyr::arrange(pgs_id)
    # Render table
    if (requireNamespace("DT", quietly = TRUE)) {
      DT::datatable(scorefile_metadata, rownames = FALSE, escape = FALSE,
                    options = list(dom = 't'))
    } else {
      knitr::kable(scorefile_metadata, format = "html", escape = FALSE)
    }
  }
} else {
  cat("**Note:** Certains scores proviennent de fichiers de score personnalisés. Pour plus d'informations, veuillez consulter la source de ces fichiers.")
}
```

Ce score a été ajusté avec le Z-score de la population la plus similaire (Z_MostSimilarPop) en utilisant [pgsc-calc](https://pgsc-calc.readthedocs.io/en/latest/) version 2.0.1 [^1].
:::


::: {.callout-tip}
## Interprétation des scores PRS

Pour chaque score polygénique évalué, le profil génétique du patient est analysé en comparaison avec un groupe de référence génétiquement diversifié. Le percentile de risque final est déterminé en alignant le score individuel avec la population de référence présentant l'ascendance génétique la plus proche.
Il est donc essentiel de se référer aux spécifications détaillées disponibles dans le Polygenic Score Catalog pour chaque score. La pertinence des estimations de risque pour d'autres populations est directement liée à leur ressemblance avec les cohortes initiales ayant servi à l'élaboration du modèle.
:::


::: {.callout-warning}
## À des fins de recherche uniquement

Une validation par un laboratoire spécialisé en génétique et l'approbation du médecin traitant sont nécessaires avant toute décision médicale ou autre basée sur ces résultats. Ces analyses constituent des données de recherche préliminaires.
:::

::: {.callout-important}
## Contraintes des PRS

Les principales restrictions comprennent une héritabilité incomplète et une connaissance limitée des facteurs de risque non génétiques. Les prédictions comportent une marge d'incertitude naturelle, et le niveau de risque peut varier selon divers éléments comme l'âge ou les traitements. Les scores de risque polygénique servent d'indicateurs prédictifs et ne constituent pas un outil diagnostique.
:::

[^1]: Lambert, S.A., Wingfield, B. et al. (2024) Enhancing the Polygenic Score Catalog with tools for score calculation and ancestry normalization. *Nature Genetics*. doi:10.1038/s41588-024-01937-x.

```{r}
#| echo: false
#| warning: false
#| message: false

# Display summary of all PRS scores for this patient
patient_summary_table <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  select(PGS, Z_MostSimilarPop, Overall_Percentile) %>%
  distinct() %>%  # remove any duplicate rows
  rename(
    "Score Type" = PGS,
    "Score" = Z_MostSimilarPop,
    "Overall Percentile" = Overall_Percentile
  )

# For HTML output, use kable
knitr::kable(patient_summary_table, format = "html", digits = 2, align = "c", 
             caption = paste("Résumé de tous les scores PRS pour le patient", params$patient_id))
```

```{r generate-sections, results='asis', echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8, out.width="100%"}
#| fig-width: 12
#| fig-height: 8
#| out-width: "100%"
#| fig-dpi: 300

# Loop through each PRS score and create a section for it
for (pgs_id in patient_pgs_scores) {
  # Get metadata for this PGS
  metadata <- get_pgs_metadata(pgs_id)
  
  # Create section header with metadata
  if (!is.null(metadata) && metadata$header$pgs_name != "") {
    cat("\n\n## Analyse du score PRS:", metadata$header$pgs_name, "\n\n")
    cat("**Identifiant PGS:**", stringr::str_replace(pgs_id, "^(PGS[0-9]+).*", "\\1"), "\n\n")
  } else {
    cat("\n\n## Analyse du score PRS:", stringr::str_replace(pgs_id, "^(PGS[0-9]+).*", "\\1"), "\n\n")
  }
  
  # NEW: Bell curve risk visualization
  cat("\n\n### Évaluation du risque génétique pour", pgs_id, "\n\n")
  
  # Create and print bell curve visualization
  bell_curve_plot <- create_bell_curve_risk_viz(params$patient_id, pgs_id)
  print(bell_curve_plot)

  # add spacing
  cat("\n\n")

  # Add risk direction text with arrow 
  cat('<div style="text-align: center; font-size: 24px; margin-top: -20px; margin-bottom: 20px;"><strong>Risque plus faible ←———→ Risque plus élevé</strong></div>\n\n')
  
  # Add explanation text
  cat("\n\n")
  cat("**Explication du score polygénique :** Votre score polygénique se situe au **", 
      round(scores_popsim %>% filter(IID == full_patient_id, PGS == pgs_id) %>% pull(Overall_Percentile)), 
      "e percentile**. Cela ne signifie pas que vous avez une probabilité de ", 
      round(scores_popsim %>% filter(IID == full_patient_id, PGS == pgs_id) %>% pull(Overall_Percentile)), 
      "% de développer cette condition. Cela signifie plutôt que sur 100 personnes, votre score polygénique est plus élevé que celui de ", 
      round(scores_popsim %>% filter(IID == full_patient_id, PGS == pgs_id) %>% pull(Overall_Percentile)), 
      " personnes et identique ou inférieur à celui de ", 
      100 - round(scores_popsim %>% filter(IID == full_patient_id, PGS == pgs_id) %>% pull(Overall_Percentile)), 
      " personnes.\n\n")
 
}
``` 

