---
title: "Analyse du risque polygénique "
format:
  html:
    toc: true  # Add table of contents for multiple scores
    embed-resources: true  # Make self-contained HTML with all resources embedded
    self-contained: true  # Ensure all resources are embedded
    fig-width: 12  # Increased from 8
    fig-height: 8  # Increased from 6
    theme: default
    code-fold: true
    code-tools: false
    fig-dpi: 300
    fig-format: "png"  # Added to ensure consistent rendering
    fig-retina: 2  # Added for better resolution

execute:
  cache: false
  freeze: auto
params:
  patient_id: "default"
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}

# Core data manipulation and visualization
library(tidyverse)  # Includes dplyr, ggplot2, etc.
library(patchwork)  # For combining plots
library(ggpubr)     # For publication-ready plots
library(viridis)    # For color scales
# library(khroma)     # For Crameri color schemes

# Text and formatting
library(showtext)   # For custom fonts
library(knitr)      # For table formatting

# Set up FontAwesome
# Check if fonts exists in the same directory as the template
use_fontawesome <- FALSE
if (file.exists("fontawesome-webfont.ttf")) {
  font_add('FontAwesome', 'fontawesome-webfont.ttf')
  showtext_auto()
  use_fontawesome <- TRUE
  message("FontAwesome font loaded successfully")
} else {
  warning("FontAwesome font not found. Using circle fallback.")
  showtext_auto()
}

# Helper function to create nicer labels
nice_label <- function(x) {
  x %>% 
    str_replace_all("_", " ") %>%
    str_to_title()
}

# Create a publication-like theme with scaled elements
publication_theme <- function(base_size = 24) {
  theme_pubr(base_size = base_size) +
    theme(
      plot.title = element_text(size = rel(1.5), face = "bold"),
      plot.subtitle = element_text(size = rel(1.2)),
      axis.title = element_text(size = rel(1.1), face = "bold"),
      axis.text = element_text(size = rel(1)),
      legend.title = element_text(size = rel(1.1)),
      legend.text = element_text(size = rel(1)),
      legend.key.size = unit(1.5, "line"),
      plot.margin = margin(t = 15, r = 15, b = 15, l = 15)
    )
}

# Get current date in a nice format
report_date <- format(Sys.Date(), "%Y-%m-%d")


```

```{r load-data, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}

# Set paths
score_files <- list.files(".", pattern = "pgs.txt", full.names = FALSE)
popsim_files <- list.files(".", pattern = "popsimilarity.txt", full.names = FALSE)
metadata_file <- list.files(".", pattern = "log_scorefiles.json", full.names = FALSE)

# Load score data
scores <- read_tsv(file.path(score_files))

# Load popsim data
popsim <- read_tsv(file.path(popsim_files))

# Load PGS metadata if available
pgs_metadata <- NULL
if (length(metadata_file) > 0) {
  tryCatch({
    pgs_metadata <- jsonlite::fromJSON(metadata_file[1], simplifyVector = FALSE)
    message("Loaded PGS metadata for ", length(pgs_metadata), " scoring files")
  }, error = function(e) {
    message("Could not load PGS metadata: ", e$message)
  })
}

# Get basic stats
n_scores <- length(unique(scores$PGS))
n_samples <- length(unique(scores$IID))

# Create comprehensive run_patient_data with all necessary variables
scores_popsim <- scores %>%
  left_join(popsim %>% select(IID, MostSimilarPop), by = "IID") %>% 
  # Create a simple ID column using the last part after underscore
  mutate(simple_id = str_extract(IID, "[^_]+$"))

# Calculate percentiles for each PRS score separately
scores_popsim <- scores_popsim %>%
  group_by(PGS) %>%
  mutate(Overall_Percentile = round(percent_rank(Z_MostSimilarPop) * 100, 1)) %>%
  group_by(PGS, MostSimilarPop) %>%
  mutate(Population_Percentile = round(percent_rank(Z_MostSimilarPop) * 100, 1)) %>%
  ungroup()

run_patient_data <- scores_popsim %>%
  filter(sampleset != "reference")

# Get reference data
reference_data <- scores_popsim %>% filter(sampleset == "reference")

# Look up full IID from simple ID
full_patient_id <- scores_popsim %>%
  filter(simple_id == params$patient_id) %>%
  pull(IID) %>%
  first()

if (is.na(full_patient_id)) {
  stop("Patient ID not found")
}

# Get all PRS scores for this patient
patient_pgs_scores <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  pull(PGS) %>%
  unique()

# Save summary data to CSV
patient_summary_all <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  select(IID, PGS, Z_MostSimilarPop, Overall_Percentile, Population_Percentile)

write_csv(patient_summary_all, "patient_summaries.csv")

# Get list of PRS scores for this patient
patient_pgs_list <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  pull(PGS) %>%
  unique()
# Normalize to plain character and trim whitespace
patient_pgs_list <- unique(trimws(as.character(patient_pgs_list)))

# Build ordered list for section generation: PGS accessions first, custom scores last
pgs_mask <- grepl('^PGS[0-9]+', patient_pgs_list, ignore.case = TRUE, perl = TRUE)
patient_pgs_scores <- c(patient_pgs_list[pgs_mask], patient_pgs_list[!pgs_mask])

# Fixed mapping between PGS IDs and French labels
pgs_french_labels <- list(
  "PGS001779" = "le syndrome de Brugada",
  "PGS002276" = "QT intervalle",
  "PGS004862" = "Cardiomyopathie dilatée",
  "PGS004911" = "Cardiomyopathie hypertrophique",
  "PGS005168" = "Fibrillation auriculaire",
  "HaydarlouHCM" = "CMH personnalisé"
)

# Function to get French label for a PGS ID
get_french_label <- function(pgs_id) {
  # Extract base ID (e.g., PGS005168 from PGS005168_hmPOS_GRCh38)
  base_id <- stringr::str_replace(pgs_id, "^(PGS[0-9]+).*", "\\1")
  # Also check for custom scores that match exactly
  if (pgs_id %in% names(pgs_french_labels)) {
    return(pgs_french_labels[[pgs_id]])
  } else if (base_id != "" && base_id %in% names(pgs_french_labels)) {
    return(pgs_french_labels[[base_id]])
  }
  # Return NULL if not in mapping (will fall back to metadata)
  return(NULL)
}

# Function to get PGS metadata
get_pgs_metadata <- function(pgs_id) {
  if (is.null(pgs_metadata)) return(NULL)
  # Normalize to base PGS accession (handles values like PGS005168_hmPOS_GRCh38)
  base_id <- stringr::str_replace(pgs_id, "^(PGS[0-9]+).*", "\\1")
  for (entry in pgs_metadata) {
    if (entry$pgs_id == base_id) {
      return(entry)
    }
  }
  return(NULL)
}
```


```{r functions, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}

# Add a direct scaling function for plots
scale_plot <- function(p) {
  p + theme(
    text = element_text(size = rel(3)),  # Scale all text by 3x
    plot.title = element_text(size = rel(3.5)),
    plot.subtitle = element_text(size = rel(3)),
    axis.title = element_text(size = rel(3)),
    axis.text = element_text(size = rel(2.5)),
    legend.text = element_text(size = rel(2.5)),
    legend.title = element_text(size = rel(3))
  )
}

# create_bell_curve_risk_viz <- function(patient_id, pgs_id = NULL) {
#   # Restore original visual but ensure we always build on a ggplot object
#   filtered_data <- if (!is.null(pgs_id)) {
#     scores_popsim %>% dplyr::filter(PGS == pgs_id)
#   } else {
#     scores_popsim
#   }

#   if (nrow(filtered_data %>% dplyr::filter(IID == full_patient_id)) == 0) {
#     return(ggplot() + annotate("text", x = 0, y = 0, label = "Patient not found") + theme_void())
#   }

#   patient_percentile <- filtered_data %>%
#     dplyr::filter(IID == full_patient_id) %>%
#     dplyr::pull(Overall_Percentile)

#   # Build bell curve on a fixed grid for stable gradient
#   x_vals <- seq(-4, 4, length.out = 2001)
#   y_vals <- dnorm(x_vals)
#   df_bell <- tibble(x = x_vals, y = y_vals)
#   max_y <- max(df_bell$y)
#   patient_x <- qnorm(patient_percentile/100)
#   # pastel line color that contrasts with all three panels
#   patient_line_color <- "#8B5CF6"  # pastel violet
  
#   # Percentile thresholds (20th and 80th)
#   t_low <- qnorm(0.20)
#   t_high <- qnorm(0.80)
  
#   # Create inverted roma color palette
#   roma_pal <- color("roma")
  
#   # Map x values to color values (x ranges from -4 to 4, map to 0-1 for palette)
#   df_bell$color_value <- (df_bell$x + 4) / 8  # normalize to 0-1
  
#   # Generate roma colors for gradient (inverted)
#   n_colors <- 100
#   roma_colors <- rev(roma_pal(n_colors))
  
#   p <- ggplot(df_bell, aes(x = x, y = y)) +
#     # Create smooth gradient using roma colors
#     geom_segment(aes(x = x, xend = x, y = 0, yend = y, color = color_value), 
#                  size = 2.5, alpha = 0.9) +
#     scale_color_gradientn(
#       colours = roma_colors,
#       guide = "none"
#     ) +
#     # Bell curve outline
#     geom_line(color = "gray70", linewidth = 3, alpha = 0.9) +
#     # Visual thresholds (using gradient colors instead)
#     # geom_vline(xintercept = c(t_low, t_high), linetype = "dashed",
#     #            linewidth = 1.2, color = c("#5DADE2", "#E57373"), alpha = 0.6) +
    
#     # Patient line (limited height to match bell curve, no arrow head)
#     annotate("segment", x = patient_x, xend = patient_x, 
#              y = 0, yend = max_y, color = patient_line_color, linewidth = 3) +
    
#     # Add human icons in each risk zone using FontAwesome - positioned very close to x-axis
#     # Low risk (left) - use light color from roma
#     annotate("text", x = -2.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
#              family = if(use_fontawesome) "FontAwesome" else "sans",
#              size = 22, color = roma_colors[1]) +  # light color from roma
#     annotate("text", x = -0.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
#              family = if(use_fontawesome) "FontAwesome" else "sans",
#              size = 18, color = "gray50") +
#     annotate("text", x = 0, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
#              family = if(use_fontawesome) "FontAwesome" else "sans",
#              size = 18, color = "gray50") +
#     annotate("text", x = 0.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
#              family = if(use_fontawesome) "FontAwesome" else "sans",
#              size = 18, color = "gray50") +
#     # High risk (right) - use dark color from roma
#     annotate("text", x = 2.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
#              family = if(use_fontawesome) "FontAwesome" else "sans",
#              size = 22, color = roma_colors[n_colors]) +  # dark color from roma
    
#     # Patient marker icon - positioned at bell curve height
#     annotate("text", x = patient_x, y = max_y, 
#              label = if(use_fontawesome) "\uf007" else "●", 
#              family = if(use_fontawesome) "FontAwesome" else "sans",
#              size = 22, color = patient_line_color) +
#     # show "Vous" next to the marker line
#     annotate("text", x = patient_x + 0.05, y = max_y*0.92, label = "Vous",
#              color = "black", size = 12, hjust = 0, vjust = 0.5) +
    
#     # Risk labels - positioned below the icons
#     annotate("text", x = -2.5, y = max_y*0.12, label = "Risque faible", size = 10) +
#     annotate("text", x = 0, y = max_y*0.12, label = "Risque moyen", size = 10) +
#     annotate("text", x = 2.5, y = max_y*0.12, label = "Risque élevé", size = 10) +
    
#     # Text below the patient marker line - larger text
#     annotate("text", x = qnorm(patient_percentile/100), y = 0.45, 
#              label = paste0("Votre score polygénique est au ", round(patient_percentile), "e percentile"), 
#              color = "black", size = 7) +
    
#     # Limit y-axis to include text and remove axis ticks
#     scale_y_continuous(limits = c(0, max_y), expand = expansion(mult = c(0, 0))) +
    
#     # no grids and axis ticks
#     theme_pubr(base_size = 36) +
#     theme(
#       panel.grid.major = element_blank(),
#       panel.grid.minor = element_blank(),
#       axis.title = element_blank(),
#       axis.text.y = element_blank(),
#       axis.ticks.y = element_blank(),
#       axis.text.x = element_blank(),
#       axis.ticks.x = element_blank(),
#       plot.title = element_text(size = 44, face = "bold", hjust = 0.5),
#       plot.subtitle = element_text(size = 32, hjust = 0.5)
#     ) +
#     labs(
#       title = paste("Score de risque polygénique :", if(!is.null(pgs_id)) pgs_id else "Global")
#     ) +
#     # Labels for risk sections - positioned outside the curve for better visibility
#     annotate("text", x = -2.75, y = max_y*0.65, 
#              label = "Moins d'individus ont un risque réduit", 
#              size = 12, color = "#2E2E2E", fontface = "bold") +
#     annotate("text", x = 0, y = max_y*0.65, 
#              label = "La plupart des individus ont un risque moyen", 
#              size = 12, color = "#2E2E2E", fontface = "bold") +
#     annotate("text", x = 2.75, y = max_y*0.65, 
#              label = "Moins d'individus ont un risque accru", 
#              size = 12, color = "#2E2E2E", fontface = "bold") +
#     # Quantile-based x-axis; keep y scale from earlier (0..0.5)
#     scale_x_continuous(
#       limits = c(-4, 4),
#       breaks = qnorm(c(0, 0.05, 0.20, 0.50, 0.80, 0.95, 0.99)),
#       labels = c("0", "5th", "20th", "50th", "80th", "95th", "99th")
#     )

#   return(p)
# }

create_bell_curve_purple_viz <- function(patient_id, pgs_id = NULL) {
  # Purple-to-red gradient version
  filtered_data <- if (!is.null(pgs_id)) {
    scores_popsim %>% dplyr::filter(PGS == pgs_id)
  } else {
    scores_popsim
  }

  if (nrow(filtered_data %>% dplyr::filter(IID == full_patient_id)) == 0) {
    return(ggplot() + annotate("text", x = 0, y = 0, label = "Patient not found") + theme_void())
  }

  patient_percentile <- filtered_data %>%
    dplyr::filter(IID == full_patient_id) %>%
    dplyr::pull(Overall_Percentile)

  # Build bell curve on a fixed grid for stable gradient
  x_vals <- seq(-4, 4, length.out = 2001)
  y_vals <- dnorm(x_vals)
  df_bell <- tibble(x = x_vals, y = y_vals)
  max_y <- max(df_bell$y)
  patient_x <- qnorm(patient_percentile/100)
  # pastel line color that contrasts with all three panels
  patient_line_color <- "#8B5CF6"  # pastel violet
  
  # Percentile thresholds (20th and 80th)
  t_low <- qnorm(0.20)
  t_high <- qnorm(0.80)
  
  p <- ggplot(df_bell, aes(x = x, y = y)) +
    # Create smooth gradient using very subtle purple progression
    # Map: left (low risk) -> almost white lavender, right (high risk) -> medium lavender
    geom_segment(aes(x = x, xend = x, y = 0, yend = y, color = x), 
                 size = 2.5, alpha = 0.9) +
    scale_color_gradientn(
      colours = c(
        "#F9F4FC",  # lightest purple - far left (low risk)
        "#F9F4FC",  # very light purple
        "#EEE0F7",  # light purple
        "#E5D4F3",  # still light (middle point)
        "#AB47BC",  # light-medium purple (just after middle, still not dark)
        "#AB47BC",  # dark purple
        "#7B1FA2"   # darkest purple (high risk)
      ), guide = "none"
    ) +
    # Bell curve outline
    geom_line(color = "gray70", linewidth = 3, alpha = 0.9) +
    # Visual thresholds (commented out - using gradient colors instead)
    # geom_vline(xintercept = c(t_low, t_high), linetype = "dashed",
    #            linewidth = 1.2, color = c("#5DADE2", "#E57373"), alpha = 0.6) +
    
    # Patient line (limited height to match bell curve, no arrow head)
    annotate("segment", x = patient_x, xend = patient_x, 
             y = 0, yend = max_y, color = patient_line_color, linewidth = 3) +
    
    # Add human icons in each risk zone using FontAwesome - positioned very close to x-axis
    # Low risk (left) - use light color
    annotate("text", x = -2.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 22, color = "#F3E5F5") +  # very light purple - low risk
    annotate("text", x = -0.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 18, color = "gray50") +
    annotate("text", x = 0, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 18, color = "gray50") +
    annotate("text", x = 0.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 18, color = "gray50") +
    # High risk (right) - use dark color
    annotate("text", x = 2.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 22, color = "#7B1FA2") +  # dark purple - high risk
    
    # Patient marker icon - positioned at bell curve height
    annotate("text", x = patient_x, y = max_y, 
             label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 22, color = patient_line_color) +
    # show "Vous" next to the marker line
    annotate("text", x = patient_x + 0.05, y = max_y*0.92, label = "Vous",
             color = "black", size = 12, hjust = 0, vjust = 0.5) +
    
    # Risk labels - positioned below the icons
    annotate("text", x = -2.5, y = max_y*0.12, label = "Risque faible", size = 10) +
    annotate("text", x = 0, y = max_y*0.12, label = "Risque moyen", size = 10) +
    annotate("text", x = 2.5, y = max_y*0.12, label = "Risque élevé", size = 10) +
    
    # Text below the patient marker line - larger text
    annotate("text", x = qnorm(patient_percentile/100), y = 0.45, 
             label = paste0("Votre score polygénique est au ", round(patient_percentile), "e percentile"), 
             color = "black", size = 7) +
    
    # Limit y-axis to include text and remove axis ticks
    scale_y_continuous(limits = c(0, max_y), expand = expansion(mult = c(0, 0))) +
    
    # Publication theme without grids and axis ticks
    theme_pubr(base_size = 36) +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.title = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      plot.title = element_text(size = 44, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 32, hjust = 0.5)
    ) +
    labs(
      title = paste("Score de risque polygénique :", if(!is.null(pgs_id)) pgs_id else "Global")
    ) +
    # Labels for risk sections - positioned outside the curve for better visibility
    annotate("text", x = -2.75, y = max_y*0.65, 
             label = "Moins d'individus ont un risque réduit", 
             size = 12, color = "#2E2E2E", fontface = "bold") +
    annotate("text", x = 0, y = max_y*0.65, 
             label = "La plupart des individus ont un risque moyen", 
             size = 12, color = "#2E2E2E", fontface = "bold") +
    annotate("text", x = 2.75, y = max_y*0.65, 
             label = "Moins d'individus ont un risque accru", 
             size = 12, color = "#2E2E2E", fontface = "bold") +
    # Quantile-based x-axis; keep y scale from earlier (0..0.5)
    scale_x_continuous(
      limits = c(-4, 4),
      breaks = qnorm(c(0, 0.05, 0.20, 0.50, 0.80, 0.95, 0.99)),
      labels = c("0", "5th", "20th", "50th", "80th", "95th", "99th")
    )

  return(p)
}
```



<!-- # Analyse du risque polygénique  -->
**Date d'émission :** `r report_date`

::: {.callout-note}
## Ce rapport contient l'analyse des scores de risque polygénique (PRS) pour le patient **`r params$patient_id`**.   

:::

::: {.callout-tip}
## Qu'est-ce qu'un score de risque polygénique?

Un score de risque polygénique est une estimation de votre risque génétique de développer une condition. Ce score est calculé à partir de nombreux  changements génétiques mineurs fréquents dans la population. Chacun de ces variants a un petit effet sur votre risque, mais tous ces effets peuvent être additionnés pour obtenir un score global, ce qui représente votre score de risque polygénique. Ce score n'est pas un diagnostic, et il ne peut pas prédire si vous développerez la condition. Il s'agit d'une estimation de votre risque comparativement au reste de la population.
:::

::: {.callout-tip}
## Ce que ce résultat signifie pour vous
*À compléter selon les besoins cliniques*; 

***V.0.1, idea:***: Pour chaque score polygénique évalué, le profil génétique du patient est analysé en comparaison avec un groupe de référence génétiquement diversifié. Le percentile de risque final est déterminé en alignant le score individuel avec la population de référence présentant l'ascendance génétique la plus proche. Il est donc essentiel de se référer aux spécifications détaillées disponibles dans le Polygenic Score Catalog pour chaque score. La pertinence des estimations de risque pour d'autres populations est directement liée à leur ressemblance avec les cohortes initiales ayant servi à l'élaboration du modèle.
:::

::: {.callout-warning}
## Limites des scores de risque polygéniques
Les scores de risque polygénique ne prennent pas en compte tous les facteurs influençant le risque de développer une condition, comme par exemple des variants génétiques rares, l'âge, les antécédents médicaux, les habitudes de vie ou l'environnement. Ils servent d'indicateurs prédictifs et ne constituent pas un outil diagnostique. Ces scores comportent une marge d'incertitude significative; ainsi, un individu avec un score élevé peut ne jamais développer la condition, tandis qu'un individu avec un score faible peut développer la condition. <br />

De plus, les scores de risque polygénique sont principalement basés sur des données génomiques d'individus d'origine ethnique européenne. Pour mieux tenir compte des différences génétiques entre différentes populations, une méthode d’ajustement basée sur l’ascendance est appliquée (PMID: 39327485). Cette correction améliore la généralisation des scores chez des individus de différentes origines, mais elle ne permet pas d'éliminer totalement le biais de généralisation. Les scores de risque polygéniques peuvent donc être moins fiables chez les personnes issues de populations non européennes.
:::

::: {.callout-caution}
## À des fins de recherche uniquement

Une validation par un laboratoire spécialisé en génétique et l'approbation du médecin traitant sont nécessaires avant toute décision médicale ou autre basée sur ces résultats. Ces analyses constituent des données de recherche préliminaires.
:::

[^1]: Lambert, S.A., Wingfield, B. et al. (2024) Enhancing the Polygenic Score Catalog with tools for score calculation and ancestry normalization. *Nature Genetics*. doi:10.1038/s41588-024-01937-x.

::: {.callout-important}
## Comment lire le graphique de risque polygénique?
Le graphiques montre où se situe votre score de risque polygénique (ligne verticale) par rapport à la population générale. La plupart des personnes ont un score situé vers le centre de la courbe, ce qui correspond à un risque génétique moyen. Moins de personnes se trouvent aux extrémités. Un score plus à gauche indique un risque plus faible que la moyenne. Un score au centre indique un risque moyen. Un score plus à droite indique un risque plus élevé que la moyenne. Les détails des scores utilisés dans ces graphiques se trouvent à la fin du Rapport.
:::

```{r generate-sections, results='asis', echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8, out.width="100%"}
#| fig-width: 12
#| fig-height: 8
#| out-width: "100%"
#| fig-dpi: 300

# Loop through each PRS score and create a section for it
for (pgs_id in patient_pgs_scores) {
  # Check fixed mapping first
  french_label <- get_french_label(pgs_id)
  
  # Get metadata for this PGS
  metadata <- get_pgs_metadata(pgs_id)
  
  # Determine display_trait: use fixed mapping if available, otherwise use metadata
  if (!is.null(french_label)) {
    display_trait <- french_label
  } else if (!is.null(metadata) && metadata$header$pgs_name != "") {
    trait_name <- metadata$header$pgs_name
    # Get the actual trait name for display
    display_trait <- if (!is.null(metadata$header$trait_reported) && metadata$header$trait_reported != "") {
      metadata$header$trait_reported
    } else if (!is.null(metadata$header$trait_mapped) && length(metadata$header$trait_mapped) > 0) {
      metadata$header$trait_mapped[[1]]
    } else {
      trait_name  # fallback to trait name
    }
  } else {
    # Custom score or no metadata - extract from ID
    if (stringr::str_detect(pgs_id, "_")) {
      display_trait <- stringr::str_replace(pgs_id, ".*_(.+)", "\\1")
    } else {
      display_trait <- params$patient_id  # fallback
    }
  }
  
  # Only extract PGSID prefix, e.g. go from PGS000001_hmPOS_GRCh38 to PGS000001
  pgs_id_section_label <- stringr::str_replace(pgs_id, "^(PGS[0-9]+).*", "\\1")
  if (pgs_id_section_label == "") {
    pgs_id_section_label <- pgs_id  # For custom scores without PGS prefix
  }
  cat("\n\n## Analyse du score PRS:", pgs_id_section_label, "pour", display_trait, "\n\n")
  
  # Bell curve risk visualization
  # cat("\n\n### Évaluation du risque génétique pour", display_trait, "\n\n")

  # Get patient data for this PGS and compute the percentile
  patient_data <- scores_popsim %>% filter(IID == full_patient_id, PGS == pgs_id)
  patient_percentile <- round(patient_data$Overall_Percentile)
  
  # Determine risk level description
  risk_level <- if (patient_percentile <= 20) {
    "plus faible que"
  } else if (patient_percentile >= 80) {
    "plus élevé que"
  } else {
    "similaire à"
  }

  # Use display_trait already determined above (uses fixed mapping or metadata)
  if (!is.null(display_trait) && display_trait != params$patient_id) {
    cat("Votre risque polygénique de développer <strong>", display_trait, "</strong> est ", risk_level, " celui de la population générale.\n\n")
  } else {
    cat("Votre risque polygénique est ", risk_level, " celui de la population générale.\n\n")
    cat("<em>Score personnalisé - veuillez contacter la source pour l'évaluation du trait PRS.</em>\n\n")
  }
  
  # Generate the summary text
  # cat("**Résumé des résultats :**\n\n")

  cat("Votre score se situe au <strong>", patient_percentile, "e percentile</strong>. Cela signifie que, comparativement à vous, ", patient_percentile, "% de la population a un risque inférieur ou égal de développer cette condition, tandis que ", 100 - patient_percentile, "% de la population a un risque supérieur. Ceci ne correspond <u>ne</u> <u>pas</u> à un risque de ", patient_percentile, "% de développer cette condition. Voici une représentation visuelle de votre risque :\n\n")
  
  # Create and print bell curve visualizations
  bell_curve_plot <- create_bell_curve_purple_viz(params$patient_id, pgs_id)
  print(bell_curve_plot)

  # add spacing
  cat("\n\n")

  # Add risk direction text with arrow - dynamically sized to match plot width with integrated arrowheads
  cat('<div style="display: flex; align-items: center; justify-content: space-between; font-size: 18px; margin-top: -20px; margin-bottom: 20px; width: 100%; padding: 0 2%;">')
  cat('<span style="white-space: nowrap;"><strong>Risque plus faible</strong></span>')
  cat('<span style="flex: 1; height: 3px; background-color: #333; margin: 0 20px; position: relative;">')
  cat('<span style="position: absolute; left: -6px; top: -4px; width: 0; height: 0; border-top: 5px solid transparent; border-bottom: 5px solid transparent; border-right: 6px solid #333;"></span>')
  cat('<span style="position: absolute; right: -6px; top: -4px; width: 0; height: 0; border-top: 5px solid transparent; border-bottom: 5px solid transparent; border-left: 6px solid #333;"></span>')
  cat('</span>')
  cat('<span style="white-space: nowrap;"><strong>Risque plus élevé</strong></span>')
  cat('</div>\n\n')
  
  # continue explanation text
  cat("\n\n")
   
}
``` 

## Résumé de tous les scores PRS pour le patient <strong>`r params$patient_id`</strong>

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: table-summary-prs

# Display summary of all PRS scores for this patient
patient_summary_table <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  select(PGS, Z_MostSimilarPop, Overall_Percentile) %>%
  distinct() %>%  # remove any duplicate rows
  mutate(.is_custom = !grepl('^PGS[0-9]+', PGS, ignore.case = TRUE)) %>%
  arrange(.is_custom, PGS) %>%
  select(-.is_custom) %>%
  rename(
    "PGS ID" = PGS,
    "Score" = Z_MostSimilarPop,
    "Overall Percentile" = Overall_Percentile
  )

# For HTML output, use kable
knitr::kable(patient_summary_table, format = "html", digits = 2, align = "c")
```

**Informations supplémentaires sur les scores de risque polygénique:** <br />

Cette analyse a été ajustée avec le Z-score de la population la plus similaire (Z_MostSimilarPop) en utilisant [pgsc-calc](https://pgsc-calc.readthedocs.io/en/latest/) version 2.0.1 [^1].

```{r, echo=FALSE, warning=FALSE}
# If metadata is available, display a compact table like the Catalog report, but
# restricted to the scores used for this patient
if (!is.null(pgs_metadata)) {
  # Keep only true PGS accessions (e.g., PGS005168) used by this patient
  patient_base_ids <- unique(stringr::str_extract(patient_pgs_list, "^(PGS[0-9]+)"))
  patient_base_ids <- patient_base_ids[!is.na(patient_base_ids) & patient_base_ids != ""]
  json_list <- pgs_metadata
  # helper extractors
  extract_chr_handle_null <- function(x, field) {
    return(replace(x[[field]], is.null(x[[field]]), ""))
  }
  link_pgscatalog <- function(id, link_type) {
    if (id != "") stringr::str_glue('<a href="https://www.pgscatalog.org/{link_type}/{id}">{id}</a>') else id
  }
  
  add_note <- function(id, note) {
    if (id != "" && note != "") {
      return(stringr::str_glue("{id} <br /> <small>{note}</small>"))
    } else {
      return(id)
    }
  }
  
  link_traits <- function(trait_efo, mapped) {
    if (length(trait_efo) == 0) {
      return("")
    } else {
      return(purrr::map2_chr(trait_efo, mapped, ~ stringr::str_glue('<a href="http://www.ebi.ac.uk/efo/{.x}">{.y}</a>')))
    }
  }
  
  extract_traits <- function(x) {
    trait_efo <- purrr::map(x, ~ extract_chr_handle_null(.x$header, "trait_efo"))
    mapped <- purrr::map(x, ~ extract_chr_handle_null(.x$header, "trait_mapped"))
    trait_display <- purrr::map2(trait_efo, mapped, link_traits)
    mapped_trait_links <- purrr::map_chr(trait_display, ~ paste(.x, collapse = "<br />"))
    reported_traits <- purrr::map(x, ~ extract_chr_handle_null(.x, "trait_reported"))
    purrr::map2_chr(reported_traits, mapped_trait_links, ~ {
      stringr::str_glue("<u>Reported trait:</u> {.x} <br /> <u>Mapped trait(s):</u> {.y}")
    })
  }
  # filter json entries to those used in this patient (Catalog-only)
  json_list_patient <- purrr::keep(json_list, ~ .x$pgs_id %in% patient_base_ids)
  if (length(json_list_patient) > 0) {
  # Create metadata table for Catalog scores
  catalog_metadata <- tibble::tibble(
    pgs_id = purrr::map_chr(json_list_patient, "pgs_id"),
    pgs_name = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "pgs_name")),
    pgp_id = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "pgp_id")),
    citation = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "citation")),
    trait_display = purrr::map2_chr(
      json_list_patient, 
      extract_traits(json_list_patient),
      ~ {
        # Check fixed mapping first
        french_label <- get_french_label(.x$pgs_id)
        if (!is.null(french_label)) {
          # Use French label from fixed mapping
          return(stringr::str_glue("<u>Trait:</u> {french_label}"))
        } else {
          # Fall back to metadata extraction
          return(.y)
        }
      }
    ),
    n_variants = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "variants_number")),
    genome_build = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "genome_build")),
    harmonised_build = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "HmPOS_build"))
  ) %>%
      dplyr::mutate(
        pgs_id = purrr::map2_chr(pgs_id, pgs_name, ~ add_note(link_pgscatalog(.x, "score"), .y)),
        pgp_id = purrr::map2_chr(pgp_id, citation, ~ add_note(link_pgscatalog(.x, "publication"), .y)),
        genome_build = purrr::map2_chr(genome_build, harmonised_build, ~ {
          if (.y != "") {
            stringr::str_glue("<u>Original build:</u> {.x} <br /> <u>Harmonised build:</u> {.y}")
          } else {
            .x
          }
        })
      )
    
    scorefile_metadata <- catalog_metadata
    
    scorefile_metadata <- scorefile_metadata %>%
      dplyr::select(pgs_id, pgp_id, trait_display, n_variants, genome_build) %>%
      dplyr::arrange(pgs_id)
    
    # Rename columns for better display
    colnames(scorefile_metadata) <- c("PGS ID", "PGP ID", "Trait", "Variants", "Build")
    
    # Render table
    if (requireNamespace("DT", quietly = TRUE)) {
      DT::datatable(scorefile_metadata, rownames = FALSE, escape = FALSE,
                    options = list(dom = 't'))
    } else {
      knitr::kable(scorefile_metadata, format = "html", escape = FALSE)
    }
  }
}

```


```{r}
#| echo: false
#| warning: false
#| message: false
#| results: asis

# Build mini-table for custom scores (those without a PGS accession)
custom_scores <- patient_pgs_list[!pgs_mask]
custom_scores <- custom_scores[!is.na(custom_scores) & custom_scores != ""]
if (length(custom_scores) > 0) {
  custom_tbl <- tibble::tibble(
    Score = custom_scores,
    Note = "Score personnalisé – consulter la source"
  )
}

if (exists("custom_tbl") && nrow(custom_tbl) > 0) {
  cat("\n\n### Scores personnalisés\n\n")
  knitr::kable(custom_tbl, format = "html", escape = FALSE)
}
```