---
title: "Analyse du risque polygénique "
format:
  html:
    toc: true  # Add table of contents for multiple scores
    embed-resources: true  # Make self-contained HTML with all resources embedded
    self-contained: true  # Ensure all resources are embedded
    fig-width: 12  # Increased from 8
    fig-height: 8  # Increased from 6
    theme: default
    code-fold: true
    code-tools: false
    fig-dpi: 300
    fig-format: "png"  # Added to ensure consistent rendering
    fig-retina: 2  # Added for better resolution
    include-in-header:
      text: |
        <style>
        .report-identifiers-footer {
          position: fixed;
          bottom: 0;
          right: 0;
          padding: 8px 16px;
          font-size: 0.85rem;
          color: #555;
          background: #f8f9fa;
          border-top: 1px solid #dee2e6;
          border-left: 1px solid #dee2e6;
          z-index: 999;
        }
        .report-identifiers-footer span { margin-left: 12px; }
        </style>
    include-after-body:
      text: |
        <div class="report-identifiers-footer" role="contentinfo">
          <span id="footer-identifier1">Identifier1</span>
          <span id="footer-identifier2">Identifier2</span>
        </div>
        <script>
        (function() {
          var p1 = document.querySelector('[data-identifier1-placeholder]');
          var p2 = document.querySelector('[data-identifier2-placeholder]');
          var f1 = document.getElementById('footer-identifier1');
          var f2 = document.getElementById('footer-identifier2');
          if (p1 && f1) f1.textContent = p1.textContent.trim();
          if (p2 && f2) f2.textContent = p2.textContent.trim();
        })();
        </script>

execute:
  cache: false
  freeze: auto
params:
  patient_id: "default"
  allowed_pgs_ids: "all"  # Comma-separated list of PGS IDs to include, or "all" for all scores
  identifier1: "Identifier1"   # Placeholder: e.g. dossier or run ID
  identifier2: "Identifier2"   # Placeholder: e.g. sample or LDM
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}

# Core data manipulation and visualization
library(tidyverse)  # Includes dplyr, ggplot2, etc.
library(patchwork)  # For combining plots
library(ggpubr)     # For publication-ready plots
library(viridis)    # For color scales
# library(khroma)     # For Crameri color schemes

# Text and formatting
library(showtext)   # For custom fonts
library(knitr)      # For table formatting

# Set up FontAwesome
# Check if fonts exists in the same directory as the template
use_fontawesome <- FALSE
if (file.exists("fontawesome-webfont.ttf")) {
  font_add('FontAwesome', 'fontawesome-webfont.ttf')
  showtext_auto()
  use_fontawesome <- TRUE
  message("FontAwesome font loaded successfully")
} else {
  warning("FontAwesome font not found. Using circle fallback.")
  showtext_auto()
}

# Helper function to create nicer labels
nice_label <- function(x) {
  x %>% 
    str_replace_all("_", " ") %>%
    str_to_title()
}

# Create a publication-like theme with scaled elements
publication_theme <- function(base_size = 24) {
  theme_pubr(base_size = base_size) +
    theme(
      plot.title = element_text(size = rel(1.5), face = "bold"),
      plot.subtitle = element_text(size = rel(1.2)),
      axis.title = element_text(size = rel(1.1), face = "bold"),
      axis.text = element_text(size = rel(1)),
      legend.title = element_text(size = rel(1.1)),
      legend.text = element_text(size = rel(1)),
      legend.key.size = unit(1.5, "line"),
      plot.margin = margin(t = 15, r = 15, b = 15, l = 15)
    )
}

# Get current date in a nice format
report_date <- format(Sys.Date(), "%Y-%m-%d")


```

```{r load-data, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}

# Set paths
score_files <- list.files(".", pattern = "pgs.txt", full.names = FALSE)
popsim_files <- list.files(".", pattern = "popsimilarity.txt", full.names = FALSE)
metadata_file <- list.files(".", pattern = "log_scorefiles.json", full.names = FALSE)

# Load score data
scores <- read_tsv(file.path(score_files))

# Load popsim data
popsim <- read_tsv(file.path(popsim_files))

# Load PGS metadata if available
pgs_metadata <- NULL
if (length(metadata_file) > 0) {
  tryCatch({
    pgs_metadata <- jsonlite::fromJSON(metadata_file[1], simplifyVector = FALSE)
    message("Loaded PGS metadata for ", length(pgs_metadata), " scoring files")
  }, error = function(e) {
    message("Could not load PGS metadata: ", e$message)
  })
}

# Get basic stats
n_scores <- length(unique(scores$PGS))
n_samples <- length(unique(scores$IID))

# Create comprehensive run_patient_data with all necessary variables
scores_popsim <- scores %>%
  left_join(popsim %>% select(IID, MostSimilarPop), by = "IID") %>% 
  # Create a simple ID column using the last part after underscore
  mutate(simple_id = str_extract(IID, "[^_]+$"))

# Look up full IID from simple ID FIRST
full_patient_id <- scores_popsim %>%
  filter(simple_id == params$patient_id) %>%
  pull(IID) %>%
  first()

if (is.na(full_patient_id)) {
  stop("Patient ID not found")
}

# Filter to reference samples + ONLY the specific target patient
# This ensures percentiles are calculated using only reference + this one patient
percentile_data <- scores_popsim %>%
  filter(sampleset == "reference" | IID == full_patient_id)

# Calculate percentiles for each PRS score separately
# Using only reference + the specific target patient
percentile_data <- percentile_data %>%
  group_by(PGS) %>%
  mutate(Overall_Percentile = round(percent_rank(Z_MostSimilarPop) * 100, 1)) %>%
  group_by(PGS, MostSimilarPop) %>%
  mutate(Population_Percentile = round(percent_rank(Z_MostSimilarPop) * 100, 1)) %>%
  ungroup()

# Update scores_popsim with correct percentiles for the patient
# Percentiles are calculated using only reference + this specific patient
patient_percentiles <- percentile_data %>%
  filter(IID == full_patient_id) %>%
  select(IID, PGS, Overall_Percentile, Population_Percentile) %>%
  rename(Overall_Percentile_correct = Overall_Percentile, Population_Percentile_correct = Population_Percentile)

# Merge patient percentiles back into scores_popsim (rename before join so _correct columns always exist)
scores_popsim <- scores_popsim %>%
  left_join(patient_percentiles, by = c("IID", "PGS"))
# Use corrected percentiles when present, else keep existing (pipeline may or may not provide percentiles)
if ("Overall_Percentile" %in% names(scores_popsim)) {
  scores_popsim <- scores_popsim %>%
    mutate(
      Overall_Percentile = coalesce(Overall_Percentile_correct, Overall_Percentile),
      Population_Percentile = coalesce(Population_Percentile_correct, Population_Percentile)
    )
} else {
  scores_popsim <- scores_popsim %>%
    mutate(
      Overall_Percentile = Overall_Percentile_correct,
      Population_Percentile = Population_Percentile_correct
    )
}
scores_popsim <- scores_popsim %>%
  select(-Overall_Percentile_correct, -Population_Percentile_correct)

run_patient_data <- scores_popsim %>%
  filter(sampleset != "reference")

# Get reference data
reference_data <- scores_popsim %>% filter(sampleset == "reference")

# Get list of PRS scores for this patient
patient_pgs_list <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  pull(PGS) %>%
  unique()
# Normalize to plain character and trim whitespace
patient_pgs_list <- unique(trimws(as.character(patient_pgs_list)))

# Filter scores based on allowed_pgs_ids parameter if provided
if (exists("params") && !is.null(params$allowed_pgs_ids) && params$allowed_pgs_ids != "" && params$allowed_pgs_ids != "all") {
  allowed_ids <- strsplit(params$allowed_pgs_ids, ",")[[1]] %>% trimws()
  
  # Function to check if a PGS ID matches any allowed ID
  # Handles both base IDs (PGS004911) and full IDs (PGS004911_hmPOS_GRCh38)
  matches_allowed <- function(pgs_id) {
    # Check exact match first (for custom scores like HaydarlouHCM)
    if (pgs_id %in% allowed_ids) {
      return(TRUE)
    }
    # Extract base ID for PGS accessions
    base_id <- stringr::str_extract(pgs_id, "^PGS[0-9]+")
    if (!is.na(base_id) && base_id %in% allowed_ids) {
      return(TRUE)
    }
    # Check if any allowed ID is a prefix of this PGS ID
    for (allowed in allowed_ids) {
      if (stringr::str_detect(pgs_id, paste0("^", stringr::str_escape(allowed)))) {
        return(TRUE)
      }
    }
    return(FALSE)
  }
  
  # Filter patient_pgs_list to only include allowed IDs
  patient_pgs_list <- patient_pgs_list[sapply(patient_pgs_list, matches_allowed)]
  
  if (length(patient_pgs_list) == 0) {
    warning("No PGS scores match the allowed IDs: ", params$allowed_pgs_ids)
  }
}

# Build ordered list for section generation: PGS accessions first, custom scores last
pgs_mask <- grepl('^PGS[0-9]+', patient_pgs_list, ignore.case = TRUE, perl = TRUE)
patient_pgs_scores <- c(patient_pgs_list[pgs_mask], patient_pgs_list[!pgs_mask])

# Save summary data to CSV
# Filter to only include scores in patient_pgs_list (which may have been filtered by allowed_pgs_ids)
patient_summary_all <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  filter(PGS %in% patient_pgs_list) %>%
  select(IID, PGS, Z_MostSimilarPop, Overall_Percentile, Population_Percentile)

write_csv(patient_summary_all, "patient_summaries.csv")

# Fixed mapping between PGS IDs and French trait names for section titles
# Format: PGS_ID -> trait name (with article: le/la)
pgs_french_traits <- list(
  "PGS001779" = "le syndrome de Brugada",
  "PGS002276" = "l'allongement de l'intervalle QT",
  "PGS004862" = "la cardiomyopathie dilatée",
  "PGS004911" = "la cardiomyopathie hypertrophique",
  "PGS005168" = "la fibrillation auriculaire",
  "HaydarlouHCM" = "la cardiomyopathie hypertrophique"
)

# Function to get French trait name for a PGS ID (for section titles)
get_french_trait <- function(pgs_id) {
  # Extract base ID (e.g., PGS005168 from PGS005168_hmPOS_GRCh38)
  base_id <- stringr::str_replace(pgs_id, "^(PGS[0-9]+).*", "\\1")
  # Also check for custom scores that match exactly
  if (pgs_id %in% names(pgs_french_traits)) {
    return(pgs_french_traits[[pgs_id]])
  } else if (base_id != "" && base_id %in% names(pgs_french_traits)) {
    return(pgs_french_traits[[base_id]])
  }
  # Return NULL if not in mapping (will fall back to metadata)
  return(NULL)
}

# Function to get French label for a PGS ID (for metadata tables - kept for backward compatibility)
get_french_label <- function(pgs_id) {
  # Use the same mapping as traits
  return(get_french_trait(pgs_id))
}

# Build indication/conditions label from the PGS shown (for preamble "Maladie/s ciblée/s")
traits_shown <- unique(na.omit(unlist(sapply(patient_pgs_scores, get_french_trait))))
indication_display <- if (length(traits_shown) > 0) paste(traits_shown, collapse = ", ") else "Tous les scores"

# Function to get PGS metadata
get_pgs_metadata <- function(pgs_id) {
  if (is.null(pgs_metadata)) return(NULL)
  # Normalize to base PGS accession (handles values like PGS005168_hmPOS_GRCh38)
  base_id <- stringr::str_replace(pgs_id, "^(PGS[0-9]+).*", "\\1")
  for (entry in pgs_metadata) {
    if (entry$pgs_id == base_id) {
      return(entry)
    }
  }
  return(NULL)
}
```


```{r functions, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}

# Add a direct scaling function for plots
scale_plot <- function(p) {
  p + theme(
    text = element_text(size = rel(3)),  # Scale all text by 3x
    plot.title = element_text(size = rel(3.5)),
    plot.subtitle = element_text(size = rel(3)),
    axis.title = element_text(size = rel(3)),
    axis.text = element_text(size = rel(2.5)),
    legend.text = element_text(size = rel(2.5)),
    legend.title = element_text(size = rel(3))
  )
}

create_bell_curve_purple_viz <- function(patient_id, pgs_id = NULL) {
  # Purple-to-red gradient version
  filtered_data <- if (!is.null(pgs_id)) {
    scores_popsim %>% dplyr::filter(PGS == pgs_id)
  } else {
    scores_popsim
  }

  if (nrow(filtered_data %>% dplyr::filter(IID == full_patient_id)) == 0) {
    return(ggplot() + annotate("text", x = 0, y = 0, label = "Patient not found") + theme_void())
  }

  patient_percentile <- filtered_data %>%
    dplyr::filter(IID == full_patient_id) %>%
    dplyr::pull(Overall_Percentile)

  # Build bell curve on a fixed grid for stable gradient
  x_vals <- seq(-4, 4, length.out = 2001)
  y_vals <- dnorm(x_vals)
  df_bell <- tibble(x = x_vals, y = y_vals)
  max_y <- max(df_bell$y)
  patient_x <- qnorm(patient_percentile/100)
  # pastel line color that contrasts with all three panels
  patient_line_color <- "#8B5CF6"  # pastel violet
  
  # Percentile thresholds (20th and 80th)
  t_low <- qnorm(0.20)
  t_high <- qnorm(0.80)
  
  p <- ggplot(df_bell, aes(x = x, y = y)) +
    # Create smooth gradient using very subtle purple progression
    # Map: left (low risk) -> almost white lavender, right (high risk) -> medium lavender
    geom_segment(aes(x = x, xend = x, y = 0, yend = y, color = x), 
                 size = 2.5, alpha = 0.9) +
    scale_color_gradientn(
      colours = c(
        "#F9F4FC",  # lightest purple - far left (low risk)
        "#F9F4FC",  # very light purple
        "#EEE0F7",  # light purple
        "#E5D4F3",  # still light (middle point)
        "#AB47BC",  # light-medium purple (just after middle, still not dark)
        "#AB47BC",  # dark purple
        "#7B1FA2"   # darkest purple (high risk)
      ), guide = "none"
    ) +
    # Bell curve outline
    geom_line(color = "gray70", linewidth = 3, alpha = 0.9) +
    # Visual thresholds (commented out - using gradient colors instead)
    # geom_vline(xintercept = c(t_low, t_high), linetype = "dashed",
    #            linewidth = 1.2, color = c("#5DADE2", "#E57373"), alpha = 0.6) +
    
    # Patient line (limited height to match bell curve, no arrow head)
    annotate("segment", x = patient_x, xend = patient_x, 
             y = 0, yend = max_y, color = patient_line_color, linewidth = 3) +
    
    # Add human icons in each risk zone using FontAwesome - positioned very close to x-axis
    # Low risk (left) - use light color
    annotate("text", x = -2.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 22, color = "#F3E5F5") +  # very light purple - low risk
    annotate("text", x = -0.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 18, color = "gray50") +
    annotate("text", x = 0, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 18, color = "gray50") +
    annotate("text", x = 0.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 18, color = "gray50") +
    # High risk (right) - use dark color
    annotate("text", x = 2.5, y = max_y*0.25, label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 22, color = "#7B1FA2") +  # dark purple - high risk
    
    # Patient marker icon - positioned at bell curve height
    annotate("text", x = patient_x, y = max_y, 
             label = if(use_fontawesome) "\uf007" else "●", 
             family = if(use_fontawesome) "FontAwesome" else "sans",
             size = 22, color = patient_line_color) +
    # show "Vous" next to the marker line
    annotate("text", x = patient_x + 0.05, y = max_y*0.92, label = "Vous",
             color = "black", size = 12, hjust = 0, vjust = 0.5) +
    
    # Risk labels - positioned below the icons
    annotate("text", x = -2.5, y = max_y*0.12, label = "Risque faible", size = 10) +
    annotate("text", x = 0, y = max_y*0.12, label = "Risque moyen", size = 10) +
    annotate("text", x = 2.5, y = max_y*0.12, label = "Risque élevé", size = 10) +
    
    # Text below the patient marker line - larger text
    annotate("text", x = qnorm(patient_percentile/100), y = 0.45, 
             label = paste0("Votre score polygénique est au ", round(patient_percentile), "e percentile"), 
             color = "black", size = 7) +
    
    # Limit y-axis to include text and remove axis ticks
    scale_y_continuous(limits = c(0, max_y), expand = expansion(mult = c(0, 0))) +
    
    # Publication theme without grids and axis ticks
    theme_pubr(base_size = 36) +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.title = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      plot.title = element_text(size = 44, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 32, hjust = 0.5)
    ) +
    labs(
      title = paste("Score de risque polygénique :", if(!is.null(pgs_id)) pgs_id else "Global")
    ) +
    # Labels for risk sections - positioned outside the curve for better visibility
    annotate("text", x = -2.75, y = max_y*0.65, 
             label = "Moins d'individus ont un risque réduit", 
             size = 12, color = "#2E2E2E", fontface = "bold") +
    annotate("text", x = 0, y = max_y*0.65, 
             label = "La plupart des individus ont un risque moyen", 
             size = 12, color = "#2E2E2E", fontface = "bold") +
    annotate("text", x = 2.75, y = max_y*0.65, 
             label = "Moins d'individus ont un risque accru", 
             size = 12, color = "#2E2E2E", fontface = "bold") +
    # Quantile-based x-axis; keep y scale from earlier (0..0.5)
    scale_x_continuous(
      limits = c(-4, 4),
      breaks = qnorm(c(0, 0.05, 0.20, 0.50, 0.80, 0.95, 0.99)),
      labels = c("0", "5th", "20th", "50th", "80th", "95th", "99th")
    )

  return(p)
}
```



<!-- # Analyse du risque polygénique  -->
<!-- **Date d'émission :** `r report_date` -->

**Identifier1:** <span data-identifier1-placeholder>`r params$identifier1`</span> · **Identifier2:** <span data-identifier2-placeholder>`r params$identifier2`</span>

::: {.callout-note}
<!-- ## Ce rapport contient l'analyse des scores de risque polygénique (PGS) pour le patient **`r params$patient_id`**.    -->

## Maladie/s ciblée/s : <strong>`r indication_display`</strong>

Ce rapport affiche uniquement les scores PGS pertinents pour ces conditions (filtrage par indication).

:::

::: {.callout-tip}
## Objectif de l'analyse

Un score de risque polygénique sert à évaluer le risque génétique d’avoir une condition, de développer cette condition et/ou des complications de cette condition. Ce score est calculé à partir de nombreux changements génétiques mineurs et qui sont fréquents dans la population. Chacun de ces variants a un petit effet sur le risque, mais tous ces effets peuvent être additionnés pour obtenir un score global, ce qui représente le score de risque polygénique. Ce score n’est pas un diagnostic, et il ne peut pas prédire avec précision si le patient développera la condition ou des complications de celle-ci. Il s’agit d’une estimation du risque comparativement au reste de la population.
:::

::: {.callout-note}
## Méthodologie

**Séquençage en parallèle à haut débit du génome entier**: ***To fill in***.

**Calcul des scores PGS (pgsc-calc)** : Les scores de risque polygénique ont été calculés avec le workflow [pgsc-calc](https://pgsc-calc.readthedocs.io/en/latest/) (Polygenic Score Catalog Calculator), qui automatise le téléchargement des fichiers de score depuis le PGS Catalog, l’appariement des variants entre ces fichiers et les échantillons cibles, et le calcul des PGS (somme linéaire des poids et dosages). Les scores présentés dans ce rapport proviennent du Polygenic Score Catalog et ont été calculés avec pgsc-calc (version 2.0.1) [^1].

**Ajustement pour l’ascendance génétique** : La distribution des PGS variant selon l’ascendance génétique, les scores ont été contextualisés à l’aide du module d’ascendance de pgsc-calc (`--run_ancestry`). Un panel de référence (par défaut 1000 Genomes + HGDP) est utilisé pour une analyse en composantes principales (ACP) ; chaque individu cible est projeté dans cet espace pour déterminer la population de référence la plus proche. Le risque relatif est rapporté en comparant le PGS de l’individu à la distribution des scores dans cette population la plus similaire : en **percentile** (colonne `percentile_MostSimilarPop`) et en **Z-score** (colonne `Z_MostSimilarPop`), comme décrit dans la [documentation pgsc-calc sur l’ascendance](https://pgsc-calc.readthedocs.io/en/latest/explanation/geneticancestry.html) [^1].

:::

::: {.callout-warning}
## Limites des scores de risque polygéniques
Les scores de risque polygénique ne prennent pas en compte tous les facteurs influençant le risque de développer une condition, comme des variants génétiques rares, l'âge, les antécédents médicaux, les habitudes de vie ou l'environnement. Ils servent d'indicateurs prédictifs et ne constituent pas un outil diagnostique. Ces scores comportent une marge d'incertitude significative; ainsi, un individu avec un score élevé peut ne jamais développer la condition, tandis qu'un individu avec un score faible peut développer la condition. <br />

De plus, la validité des estimations de risque est directement liée à la ressemblance avec les cohortes initiales ayant servi à l’élaboration du score. Par exemple, un score polygénique développé dans une population d’origine ancestrale européenne aura une performance plus élevée chez les individus d’origine ancestrale européenne que des individus d’autres origines. Au fait, les scores de risque polygénique sont principalement basés sur des données génomiques d'individus d'origine ethnique européenne. Pour mieux tenir compte des différences génétiques entre différentes populations, une méthode d'ajustement basée sur l'ascendance est appliquée (PMID: 39327485). Cette correction améliore la généralisation des scores chez des individus de différentes origines, mais elle ne permet pas d'éliminer totalement le biais de généralisation. Les scores de risque polygéniques peuvent donc être moins fiables chez les personnes issues de populations non européennes.
:::

[^1]: Lambert, S.A., Wingfield, B. et al. (2024) Enhancing the Polygenic Score Catalog with tools for score calculation and ancestry normalization. *Nature Genetics*. doi:10.1038/s41588-024-01937-x.

::: {.callout-important}
## Comment lire le graphique de risque polygénique?
Le graphiques montre où se situe le score de risque polygénique du patient (ligne verticale) par rapport à la population générale ajustée pour l’origine ancestrale. La plupart des personnes ont un score situé vers le centre de la courbe, ce qui correspond à un risque génétique moyen. Moins de personnes se trouvent aux extrémités. Un score plus à gauche indique un risque plus faible que la moyenne. Un score au centre indique un risque moyen. Un score plus à droite indique un risque plus élevé que la moyenne. Les détails des scores utilisés dans ces graphiques se trouvent à la fin du Rapport.
:::

```{r generate-sections, results='asis', echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8, out.width="100%"}
#| fig-width: 12
#| fig-height: 8
#| out-width: "100%"
#| fig-dpi: 300

# Loop through each PRS score and create a section for it
for (pgs_id in patient_pgs_scores) {
  # Extract PGS ID for section label (base ID or full custom score name)
  pgs_id_section_label <- stringr::str_replace(pgs_id, "^(PGS[0-9]+).*", "\\1")
  if (pgs_id_section_label == "") {
    pgs_id_section_label <- pgs_id  # For custom scores without PGS prefix
  }
  
  # Get French trait name from dictionary
  french_trait <- get_french_trait(pgs_id)
  
  # Determine display_trait: use dictionary first, then fall back to metadata
  if (!is.null(french_trait)) {
    display_trait <- french_trait
  } else {
    # Fall back to metadata if not in dictionary
    metadata <- get_pgs_metadata(pgs_id)
    if (!is.null(metadata) && metadata$header$pgs_name != "") {
      trait_name <- metadata$header$pgs_name
      # Get the actual trait name for display
      display_trait <- if (!is.null(metadata$header$trait_reported) && metadata$header$trait_reported != "") {
        metadata$header$trait_reported
      } else if (!is.null(metadata$header$trait_mapped) && length(metadata$header$trait_mapped) > 0) {
        metadata$header$trait_mapped[[1]]
      } else {
        trait_name  # fallback to trait name
      }
    } else {
      # Custom score or no metadata - extract from ID
      if (stringr::str_detect(pgs_id, "_")) {
        display_trait <- stringr::str_replace(pgs_id, ".*_(.+)", "\\1")
      } else {
        display_trait <- pgs_id  # fallback to PGS ID itself
      }
    }
  }
  
  # Generate section title using dictionary format: "Analyse du score (PGS_ID) pour [trait]"
  cat("\n\n## Analyse du score (", pgs_id_section_label, ") pour ", display_trait, "\n\n", sep = "")
  
  # Bell curve risk visualization
  # cat("\n\n### Évaluation du risque génétique pour", display_trait, "\n\n")

  # Get patient data for this PGS and compute the percentile
  patient_data <- scores_popsim %>% filter(IID == full_patient_id, PGS == pgs_id)
  patient_percentile <- round(patient_data$Overall_Percentile)
  
  # Determine risk level description
  risk_level <- if (patient_percentile <= 20) {
    "plus faible que"
  } else if (patient_percentile >= 80) {
    "plus élevé que"
  } else {
    "similaire à"
  }

  # Use display_trait already determined above (uses fixed mapping or metadata)
  # Remove article (le/la/l') from display_trait for body text to avoid duplication
  trait_for_text <- display_trait
  if (stringr::str_detect(display_trait, "^le |^la |^l'")) {
    trait_for_text <- stringr::str_replace(display_trait, "^(le |la |l')", "")
  }
  
  if (!is.null(display_trait) && display_trait != params$patient_id && !is.null(french_trait)) {
    # Use the trait with article from dictionary
    cat("Le risque polygénique d'avoir <strong>", display_trait, "</strong> est ", risk_level, " celui de la population générale.\n\n", sep = "")
  } else if (!is.null(display_trait) && display_trait != params$patient_id) {
    # Fallback: use trait without article
    cat("Le risque polygénique d'avoir <strong>", trait_for_text, "</strong> est ", risk_level, " celui de la population générale.\n\n", sep = "")
  } else {
    cat("Le risque polygénique est ", risk_level, " celui de la population générale.\n\n", sep = "")
    cat("<em>Score personnalisé - veuillez contacter la source pour l'évaluation du trait PRS.</em>\n\n")
  }
  
  # Generate the summary text
  # cat("**Résumé des résultats :**\n\n")

  cat("Le score se situe au <strong>", patient_percentile, "e percentile</strong>. Cela signifie que, ", patient_percentile, "% de la population a un risque inférieur ou égal d'avoir cette condition, tandis que ", 100 - patient_percentile, "% de la population a un risque supérieur. Ceci ne correspond <u>pas</u> à un risque de ", patient_percentile, "% de développer cette condition. Voici une représentation visuelle du risque pour faciliter la communication du résultat avec le patient:\n\n")
  
  # Create and print bell curve visualizations
  bell_curve_plot <- create_bell_curve_purple_viz(params$patient_id, pgs_id)
  print(bell_curve_plot)

  # add spacing
  cat("\n\n")

  # Add risk direction text with arrow - dynamically sized to match plot width with integrated arrowheads
  cat('<div style="display: flex; align-items: center; justify-content: space-between; font-size: 18px; margin-top: -20px; margin-bottom: 20px; width: 100%; padding: 0 2%;">')
  cat('<span style="white-space: nowrap;"><strong>Risque plus faible</strong></span>')
  cat('<span style="flex: 1; height: 3px; background-color: #333; margin: 0 20px; position: relative;">')
  cat('<span style="position: absolute; left: -6px; top: -4px; width: 0; height: 0; border-top: 5px solid transparent; border-bottom: 5px solid transparent; border-right: 6px solid #333;"></span>')
  cat('<span style="position: absolute; right: -6px; top: -4px; width: 0; height: 0; border-top: 5px solid transparent; border-bottom: 5px solid transparent; border-left: 6px solid #333;"></span>')
  cat('</span>')
  cat('<span style="white-space: nowrap;"><strong>Risque plus élevé</strong></span>')
  cat('</div>\n\n')
  
  # continue explanation text
  cat("\n\n")
   
}
``` 

## Résumé de tous les scores de risque polygénique pour le patient <strong>`r params$patient_id`</strong>

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: table-summary-prs

# Display summary of all PRS scores for this patient
# Filter to only include scores in patient_pgs_list (which may have been filtered by allowed_pgs_ids)
patient_summary_table <- scores_popsim %>%
  filter(IID == full_patient_id) %>%
  filter(PGS %in% patient_pgs_list) %>%
  select(PGS, Z_MostSimilarPop, Overall_Percentile) %>%
  distinct() %>%  # remove any duplicate rows
  mutate(
    # Add Condition column using dictionary mapping
    Condition = sapply(PGS, function(pgs_id) {
      trait <- get_french_trait(pgs_id)
      if (!is.null(trait)) {
        # Remove article (le/la/l') for table display
        stringr::str_replace(trait, "^(le |la |l')", "")
      } else {
        # Fall back to metadata or PGS ID
        metadata <- get_pgs_metadata(pgs_id)
        if (!is.null(metadata) && !is.null(metadata$header$trait_reported) && metadata$header$trait_reported != "") {
          metadata$header$trait_reported
        } else if (!is.null(metadata) && !is.null(metadata$header$trait_mapped) && length(metadata$header$trait_mapped) > 0) {
          metadata$header$trait_mapped[[1]]
        } else {
          pgs_id  # fallback to PGS ID
        }
      }
    }),
    .is_custom = !grepl('^PGS[0-9]+', PGS, ignore.case = TRUE)
  ) %>%
  arrange(.is_custom, PGS) %>%
  select(PGS, Condition, Z_MostSimilarPop, Overall_Percentile) %>%
  rename(
    "PGS ID" = PGS,
    "Condition" = Condition,
    "Score" = Z_MostSimilarPop,
    "Percentile" = Overall_Percentile
  )

# For HTML output, use kable
knitr::kable(patient_summary_table, format = "html", digits = 2, align = "c")
```

**Informations supplémentaires sur les scores de risque polygénique:** <br />

```{r, echo=FALSE, warning=FALSE}
# If metadata is available, display a compact table like the Catalog report, but
# restricted to the scores used for this patient
if (!is.null(pgs_metadata)) {
  # Keep only true PGS accessions (e.g., PGS005168) used by this patient
  patient_base_ids <- unique(stringr::str_extract(patient_pgs_list, "^(PGS[0-9]+)"))
  patient_base_ids <- patient_base_ids[!is.na(patient_base_ids) & patient_base_ids != ""]
  json_list <- pgs_metadata
  # helper extractors
  extract_chr_handle_null <- function(x, field) {
    return(replace(x[[field]], is.null(x[[field]]), ""))
  }
  link_pgscatalog <- function(id, link_type) {
    if (id != "") stringr::str_glue('<a href="https://www.pgscatalog.org/{link_type}/{id}">{id}</a>') else id
  }
  
  add_note <- function(id, note) {
    if (id != "" && note != "") {
      return(stringr::str_glue("{id} <br /> <small>{note}</small>"))
    } else {
      return(id)
    }
  }
  
  link_traits <- function(trait_efo, mapped) {
    if (length(trait_efo) == 0) {
      return("")
    } else {
      return(purrr::map2_chr(trait_efo, mapped, ~ stringr::str_glue('<a href="http://www.ebi.ac.uk/efo/{.x}">{.y}</a>')))
    }
  }
  
  extract_traits <- function(x) {
    trait_efo <- purrr::map(x, ~ extract_chr_handle_null(.x$header, "trait_efo"))
    mapped <- purrr::map(x, ~ extract_chr_handle_null(.x$header, "trait_mapped"))
    trait_display <- purrr::map2(trait_efo, mapped, link_traits)
    mapped_trait_links <- purrr::map_chr(trait_display, ~ paste(.x, collapse = "<br />"))
    reported_traits <- purrr::map(x, ~ extract_chr_handle_null(.x, "trait_reported"))
    purrr::map2_chr(reported_traits, mapped_trait_links, ~ {
      stringr::str_glue("<u>Reported trait:</u> {.x} <br /> <u>Mapped trait(s):</u> {.y}")
    })
  }
  # filter json entries to those used in this patient (Catalog-only)
  json_list_patient <- purrr::keep(json_list, ~ .x$pgs_id %in% patient_base_ids)
  if (length(json_list_patient) > 0) {
  # Create metadata table for Catalog scores
  catalog_metadata <- tibble::tibble(
    pgs_id = purrr::map_chr(json_list_patient, "pgs_id"),
    pgs_name = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "pgs_name")),
    pgp_id = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "pgp_id")),
    citation = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "citation")),
    trait_display = purrr::map2_chr(
      json_list_patient, 
      extract_traits(json_list_patient),
      ~ {
        # Check fixed mapping first
        french_label <- get_french_label(.x$pgs_id)
        if (!is.null(french_label)) {
          # Use French label from fixed mapping
          return(stringr::str_glue("<u>Trait:</u> {french_label}"))
        } else {
          # Fall back to metadata extraction
          return(.y)
        }
      }
    ),
    n_variants = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "variants_number")),
    genome_build = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "genome_build")),
    harmonised_build = purrr::map_chr(json_list_patient, ~ extract_chr_handle_null(.x$header, "HmPOS_build"))
  ) %>%
      dplyr::mutate(
        pgs_id = purrr::map2_chr(pgs_id, pgs_name, ~ add_note(link_pgscatalog(.x, "score"), .y)),
        pgp_id = purrr::map2_chr(pgp_id, citation, ~ add_note(link_pgscatalog(.x, "publication"), .y)),
        genome_build = purrr::map2_chr(genome_build, harmonised_build, ~ {
          if (.y != "") {
            stringr::str_glue("<u>Original build:</u> {.x} <br /> <u>Harmonised build:</u> {.y}")
          } else {
            .x
          }
        })
      )
    
    scorefile_metadata <- catalog_metadata
    
    scorefile_metadata <- scorefile_metadata %>%
      dplyr::select(pgs_id, pgp_id, trait_display, n_variants, genome_build) %>%
      dplyr::arrange(pgs_id)
    
    # Rename columns for better display
    colnames(scorefile_metadata) <- c("PGS ID", "PGP ID", "Trait", "Variants", "Build")
    
    # Render table
    if (requireNamespace("DT", quietly = TRUE)) {
      DT::datatable(scorefile_metadata, rownames = FALSE, escape = FALSE,
                    options = list(dom = 't'))
    } else {
      knitr::kable(scorefile_metadata, format = "html", escape = FALSE)
    }
  }
}

```


```{r}
#| echo: false
#| eval: false
#| warning: false
#| message: false
#| results: asis

# Build mini-table for custom scores (those without a PGS accession)
custom_scores <- patient_pgs_list[!pgs_mask]
custom_scores <- custom_scores[!is.na(custom_scores) & custom_scores != ""]
if (length(custom_scores) > 0) {
  custom_tbl <- tibble::tibble(
    Score = custom_scores,
    Note = "Score personnalisé – consulter la source"
  )
}

if (exists("custom_tbl") && nrow(custom_tbl) > 0) {
  cat("\n\n### Scores personnalisés\n\n")
  knitr::kable(custom_tbl, format = "html", escape = FALSE)
}
```

